---
author: 
  "Jeanie Na"
date: "`r format(Sys.time(), '%d %B %Y')`"
title    : "High throughput study analysis one experiment investigation"


output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
    theme: united
    highlight: tango

params:

  Gene:
    label: "Gene:"
    value: "BRCA2"
    input: text 
  Output_dir:
    label: "Output Dir:"
    value: "/respath/03_21_2024/xexon_debug/"
    input: text    
       
  Exons:
    label: "Exons included:"
    value: "15_set1,16,17,18C_bs,18N_bs,19,20,21_bs,22_set1,23,24,25C,25N,26_p3"
    input: text   
      
  loessSubset:
    label: "Loess Filter (keep):"
    value: "ModFindlay"
    choices: ["Findlay", "ModFindlay", "ModFindlay2", "Synonymous", "allSNV"]
    input: select
    
  EventCount_Filter:
    label: "EventCount Filter:"
    value: "libD510"
    choices: ["lib10", "libD510", "lib100", "libD5100"]
    input: select
  vset:
    label: "Variant set:"
    value: "SNV"
    choices: ["allv", "SNV"]
    input: select

  Score_method:
    label: "Score Method:"
    value: "B"
    choices: ["A", "B", "C", "D"]
    input: select
  fcs:
    label: "filter FC:"
    value: 1
    choices: [1, 1.5]
    input: select
  Lab_curation:
    label: "Lab Curattion:"
    value: "No"
    choices: ["Yes", "No"]
    input: select
  Ratio:
    label: "Ratio:"
    value: "D14vLib"
    choices: ["D14vLib", "D14vD5R", "D14vD5L"]
    input: select
  MM:
    label: "Mixture Model:"
    value: "normal"
    choices: ["Robust", "normal"]
    input: select
  Outlierremove:
    label: "Outlierremove Method:"
    value: "No"
    choices: ["p1", "p2", "p5", "No"]
    input: select
  extraFilter:
    label: "Supplementary filter:"
    value: "Yes"
    choices: ["Yes", "No"]
    input: select
  inputfile:
    label: "Input files:"
    value: "/respath/03_21_2024/xexon_inputfiles/BRCA2_SNV_15_set1,16,17,18C_bs,18N_bs,19,20,21_bs,22_set1,23,24,25C,25N,26_p3_D14vLib_ModFindlay_libD510_1.csv" 
    input: text
  vexclude:
    label: "variable exlusion files:"
    value: "/datapath/10_2023_curation/Varaints to be excluded when use MAVE internal control to build model.xlsx" 
    choices: ["/datapath/10_2023_curation/Varaints to be excluded when use MAVE internal control to build model.xlsx",NULL] 
    input: select
  clinvarfile:
    label: "Clinvar File:"
    value: "/datapath/10_2023_curation/ClinVar_BRCA2_DBD_control_111523_nonsense_syn_missense_exclude_single.txt"
    choices: ["/datapath/RAD51C_E2c&3_ClinVar Controls_012224.txt", "/datapath/10_2023_curation/ClinVar_BRCA2_DBD_control_111523_nonsense_syn.txt"]
    input: select
  HDRfile:
    label: "HDR File:"
    value: "/datapath/HDR_exclSPLICEhypo.xlsx"
    choices: ["/datapath/HDR_exclSPLICEhypo.xlsx", NULL]
    input: select
  targetfile:
    label: "Data input file:"
    value: "/datapath/chunling_target_ind_BRCA2.csv"
    choices: ["/datapath/chunling_target_ind_BRCA2.csv", "/datapath/10_2023_curation//chunling_target.csv", NULL] 
    input: select
  funR:
    label: "functions.R:"
    value: "./functions.R"
---

```{=html}

<style type="text/css">
.main-container {
  max-width: 1200px;
  margin-left: auto;
  margin-right: auto;
}
</style>
```
<link rel="stylesheet" type="text/css" href="http://cdn.datatables.net/1.10.5/css/jquery.dataTables.min.css">

```{=html}
<script type="text/javascript">
  $(document).ready(function() {
    $("table").DataTable();
  } );
</script>
```

```{r setup, include=TRUE, warning=FALSE, message=FALSE}
    
knitr::opts_chunk$set(warning = FALSE, echo = TRUE)
library(kableExtra)
library(plotly)
library(arsenal)
library(boot)

library(ggplot2)
library(dplyr)
library(tidyr)
library(readr)
library(readxl)
library(tidyverse)
library(threejs)
library(pROC)

source(params$funR)
resdir <- params$Output_dir
Score_method <- params$Score_method
Lab_curation <- params$Lab_curation
vset <- params$vset
Exons <- params$Exons
Ratio <- params$Ratio
Gene <- params$Gene
extraFilter <- params$extraFilter
inputfile <- params$inputfile
fc <- params$fcs
HDRfile <- params$HDRfile
```

## Setting of this report

Gene: `r  Gene`\
Inputfile: `r params$inputfile`
Exons included : `r Exons`\
Outputdir: `r  resdir`\
loessSubset:  `r params$loessSubset`\
EventCount_Filter: `r params$EventCount_Filter`\
vset: `r vset`\
Score_method: `r Score_method`\
Lab_curation: `r Lab_curation`\
Ratio: `r Ratio`\
Mixture Model: `r params$MM`\
Outlierremove: `r params$Outlierremove`\
clinvarfile: `r params$clinvarfile` \
extraFilter:  `r params$extraFilter` \
Varaints_to_be_excluded_when_use_MAVE_internal_control_to_build_model: `r params$vexclude` \

## Aim

Apply Findlay methods on functional scores. `r Ratio`.

```{r}
if(!is.null(params$vexclude)){
  Varaints_to_be_excluded_when_use_MAVE_internal_control_to_build_model <- read_excel(params$vexclude)

  Varaints_to_be_excluded_when_use_MAVE_internal_control_to_build_model <- Varaints_to_be_excluded_when_use_MAVE_internal_control_to_build_model %>% mutate(uPOS = paste(GRCh38Location-1, REF, ALT, sep = "_") )

}
                    
```

```{r}
if(!is.null(params$HDRfile)){
  if (grepl("\\.txt$", params$HDRfile)) {
    # Read the file using read.table
    HDR <- read.table(file = params$HDRfile, header = TRUE, sep = '\t')
  } else {
    HDR <- read_excel(params$HDRfile)
  }
  HDR <- HDR %>% unique()
  
  HDR$htype <- HDR$`HDR function`
}

```

```{r readin, include=TRUE, warning=FALSE, message=FALSE}
## read in data

inputfiles <-  read_csv(params$inputfile)
exons <- inputfiles$exon
dt <- list()
for(i in 1:length(exons)){
  dti <- read_csv(paste0(inputfiles$dtdir[i], inputfiles$filenames[i]), num_threads = 1)
  dt[[i]] <- dti 
}

```

## Data

This analysis does `r ifelse(Lab_curation == "Yes", "", "not" )` take Lab curation into account. 

## Analysis Plan:

0. Start with the adjusted ratio from each individual exon exploration. 

1. run the pipeline with normalization within each exon and across exons including all lab curated and cleaned data files. Files used are `r inputfiles$dtdir`

`r inputfiles$filenames`

Reference Method

589 Normalizing scores within and across exons  

590 Position-adjusted log2 day 11 over library ratios were normalized first across exon

591 replicates, and then across all exons assayed. Scores from within each replicate were linearly 

592 scaled such that the median synonymous and median nonsense SNVs within the replicate were 

593 set to the median synonymous and median nonsense SNV values averaged across replicate 

594 experiments. The ensuing SNV scores for each replicate were then normalized across exons in 

595 the same way by again using median synonymous and median nonsense SNVs.


A. Within each exon:

Positional adjusted D14vLib or D14vD5 ratios from within each replicate were linearly scaled such that the median synonymous and median nonsense SNVs within the replicate were set to the median synonymous and median nonsense SNV values averaged across replicate  experiments.

B. Normalization across exons.

The ensuing SNV scores for each replicate were then normalized across exons in  the same way by again using median synonymous and median nonsense SNVs.

C. functional score

Get the average functional score cross replicates.

2. Then calculate both 95% confidence interval (CI) and 99% CI for all stop-gain variants based on the clean data (D8/D5 and D8_olaparib/D5); 95% CI and 99% CI for all synonymous variants based on the clean data (D8/D5 and D8_olaparib/D5). For missense variants, no CI calculation needed, only functional scores from D8/D5 and D8_olaparib/D5. Classify hypomorphs based on the 95% & 99% CI from stop-gains and synonymous.

3. Calculate the p-values for the pairwise mean difference for the 3 categories as a group (between stop-gain vs hypomorphs, synonymous vs hypomorphs, stop-gain vs synonymous) for the no drug experiment and olaparib experiment. Then, once the pairwise mean differences are calculated within no-drug or olaparib experiment, compare the pairwise mean difference between no-drug experiment and olaparib experiment. 

4 Added the mixture model part, also the posterior probability 1-99% 2-98% corresponding functional score. 

```{r prepdt1}

### this step is to create necessary columns
## up to 6 replicates 
## take care of this if more replicates
getnreps <- function(names){
  n <- ifelse(length(grep("R6",names))>0, 6,
              ifelse(length(grep("R5",names))>0, 5,
                     ifelse(length(grep("R4",names))>0, 4, 
                        ifelse(length(grep("R3",names))>0, 3, 
                            ifelse(length(grep("R2",names))>0, 2, 1)))))
  n
}

nmax <- 0
for(ei in 1:length(exons)){
  dt[[ei]] <- dt[[ei]] %>% mutate(uPOS = paste(POS, REF, ALT, sep = "_"), Exon = exons[ei], AApos=as.numeric(AApos))
  ## find out how many replicates
  nr <- getnreps(colnames(dt[[ei]]))
  if(nr > nmax){
    nmax <- nr
  }
  ## create R1:Rn,  Ri_raw, Ri_loess from the right columns
  for(i in 1:nr){
    Ri <- paste0("R", i)
    Ri_raw <- paste0("R", i, "_raw")
    Ri_loess <- paste0("R", i, "_loess")
      
    if(Ratio == "D14vLib"){
      Ri_ratio <- paste0("R", i, "_D14_lib")
      ratio_adjusted_Ri <- paste0("D14_lib_ratio_adjusted_R", i) # use loess adjusted values as starting values
    }else if(Ratio == "D14vD5L"){
      Ri_ratio <- paste0("R", i, "_D14_D5")
      ratio_adjusted_Ri <- paste0("D14_D5_ratio_adjusted_R", i) # use loess adjusted values as starting values
    }else{
      Ri_ratio <- paste0("R", i, "_D14_D5")
      ratio_adjusted_Ri <- paste0("R", i, "_D14_D5") # use raw values as starting values
    }
    
    dt[[ei]] <- dt[[ei]] %>% mutate(!!sym(Ri) := as.numeric(!!sym(ratio_adjusted_Ri))) %>% 
        mutate(!!sym(Ri_raw) := as.numeric(!!sym(Ri_ratio)))%>% 
        mutate(!!sym(Ri_loess) := as.numeric(!!sym(Ri)))
  }
  ## if there is lab curation do it here.
  if(Lab_curation == "Yes"){ ## remove lab says exclude all after combination
    if(Score_method == "A"){
      dt[[ei]] <- dt[[ei]] %>% filter(!`lab CV A` %in% c("exclude all"))
    }else if(Score_method == "B"){
      dt[[ei]] <- dt[[ei]] %>% filter(!`lab CV B` %in% c("exclude all"))
    }else if(Score_method == "C"){
      dt[[ei]] <- dt[[ei]] %>% filter(!`lab CV C` %in% c("exclude all"))
    }else{
      dt[[ei]] <- dt[[ei]] %>% filter(!`lab CV D` %in% c("exclude all"))
    }
  }
}
 
```


```{r prepdt2}
## this step is to apply filtering based on score summarized, and apply extra Filter
## first create the function to apply filtering for different score and different number of replicates
## note: if there were only two reps, the name of the removeA, C, D columns are different
## 

prefilter <- function(data,extraFilter,Score_method = "A" ){
  if(Score_method == "B"){
    return(data)
  }else{
    nr <- getnreps(colnames(data))
    ## create removes from the right columns
    if(nr <3){
      removecln <-  paste0("remove", Score_method)
      data <- data %>% filter(!!sym(removecln) == FALSE)
    }else{
      for(i in 1:nr){
        removeclni <- paste0("remove", Score_method,i)
        Ri <- paste0("R",i)
        if(length(which(data[, removeclni] ==TRUE)) > 0){
          data[which(data[, removeclni] ==TRUE), Ri] <- NA
        }
      }
    }
    
    if(extraFilter == "Yes" & Score_method == "A"){
      ## remove variants if the mean differ from B if it is scoreA
      data <- data %>% filter(abs(functional.scoreA-functional.scoreB) < fc)
    }else if(extraFilter == "Yes" & Score_method == "C"){
      data <- data %>% filter(abs(functional.scoreC-functional.scoreB) < fc)
    }else{
      data <- data
    }
    return(data)
  }
}

for(ei in 1:length(exons)){
  dt[[ei]] <- prefilter(dt[[ei]],extraFilter,Score_method)
}
```


```{r prepdt3}
## this step to gather the necessary information for output
## Combine exons.
colns <- c("Exon", "uPOS", "POS", "REF",                       
"ALT", "AApos","AltAA", "AltCodon", "EventType", #"EventType2",  
"RefAA", "RefCodon", "SpliceAI_ALLELE", "SpliceAI_DP_AG", "SpliceAI_DP_AL", "SpliceAI_DP_DG",  "SpliceAI_DP_DL", "SpliceAI_DS_AG",  "SpliceAI_DS_AL", "SpliceAI_DS_DG", "SpliceAI_DS_DL", "spliceR",  "R1", "R2", "R3", "R4", "R5", "R6", "functional.scoreA", "functional.scoreB",    "functional.scoreC", "functional.scoreD")

## take care of this if number of replicates is more than 6.
rawcolns <- c("uPOS", "Exon", "R1_lib", "R1_D5", "R1_D14", "R2_lib", "R2_D5","R2_D14", "R3_lib", "R3_D5",	"R3_D14",
               "R4_lib", "R4_D5", "R4_D14",		"R5_lib", "R5_D5","R5_D14", "R6_lib", "R6_D5",	"R6_D14", 
              "R1_raw", "R2_raw", "R3_raw", "R4_raw", "R5_raw", "R6_raw",
              "R1_loess", "R2_loess", "R3_loess", "R4_loess",  "R5_loess", "R6_loess")

rawcolns_num <- setdiff(rawcolns, c("uPOS", "Exon"))

## subset to these columns only
dt0 <- dt
for(i in 1:length(exons)){
  temp <- dt[[i]]
  temp[,rawcolns_num[rawcolns_num %in% colnames(temp)]] <- apply(temp[, rawcolns_num[rawcolns_num %in% colnames(temp)]], 2, function(x) as.numeric(as.character(x))) ## make sure all these are numeric
  dt0[[i]] <- temp
  
  dt0[[i]] <- temp[, rawcolns[rawcolns %in% colnames(temp)]]
  dt[[i]] <- temp[, colns[colns %in% colnames(temp)]]
}


if(length(exons) > 1){
  dt0 <- dplyr::bind_rows(dt0)
  dt <- dplyr::bind_rows(dt)
}else{
  dt0 <- dt0[[1]]
  dt <- dt[[1]]
}
```


```{r prepdt4}
## this step is to get the long format so that we can re-normalize within exon using score0
## take care of this if number of replicates > 6
if(nmax == 6){
  dt.long <- tidyr::gather(dt, key = "Experiment", value = "score0", R1, R2, R3, R4, R5, R6)
}else if(nmax == 5){
  dt.long <- tidyr::gather(dt, key = "Experiment", value = "score0", R1, R2, R3, R4, R5)
} else if(nmax == 4){
  dt.long <- tidyr::gather(dt, key = "Experiment", value = "score0", R1, R2, R3, R4)
}else if(nmax == 3){ ## 3
  dt.long <- tidyr::gather(dt, key = "Experiment", value = "score0", R1, R2, R3)
}else if(nmax == 2){ ## 2
  dt.long <- tidyr::gather(dt, key = "Experiment", value = "score0", R1, R2)
}else{ ## 1
  dt.long <- tidyr::gather(dt, key = "Experiment", value = "score0", R1)
}

dt.long$Experiment <- as.numeric(substr(dt.long$Experiment, 2, 2))
dt.long$score0 <- as.numeric(dt.long$score0)

mydt <- dt.long 

```

```{r prepdt5}
## this step to add in Target information
## within target region
target <- read.csv(params$targetfile)
## previously for BRCA2 target$Exon <- gsub("_sg.*_*", "", target$Exon), check if the target file need to be updated.
target <- target %>% unique()

mydt <- merge(mydt, target) %>% mutate(inTarget = ifelse(POS <= End & POS >= Start, TRUE, FALSE))

```


```{r normfuns, include=TRUE}
## this step is to normalize score0 now to score1 (within exon), to score2 (xexon)
##a = (Syn.Avg.med-StopGain.Avg.med)/(Syn.medi-StopGain.medi),b =Syn.Avg.med – a*Syn.medi 

myscale <- function(x, Syn.Avg.med, StopGain.Avg.med, Syn.med1, StopGain.med1){
  a <- (Syn.Avg.med-StopGain.Avg.med)/(Syn.med1-StopGain.med1)
  b  <- Syn.Avg.med - a*Syn.med1
  x2 <- a*x +b
  return(x2)
}

norm_win_exon <- function(dt){ ## note: remove splice variants for calculation medians inside so to keep the splice variants later in the normalization.
  dt.out <- NULL
  
  for (ei in unique(dt$Exon)){
    dti <- dt %>% filter(Exon == ei)
    dti.temp <- dti %>% ## remove variants in splicing region
      filter(spliceR == FALSE) %>% filter(nchar(REF) == 1 & nchar(ALT) == 1) %>%##filter(REF %in% c("A", "T", "C", "G") & ALT %in% c("A", "T", "C", "G") )%>% 
      filter(inTarget == TRUE) 
    if(!is.null(params$vexclude)){
      dti.temp <- dti.temp %>% filter(!uPOS %in% Varaints_to_be_excluded_when_use_MAVE_internal_control_to_build_model$uPOS)  ## remove variants on chunling's list
    }
      
    meds <- dti.temp %>% group_by(EventType, Experiment) %>%
      dplyr::summarize(M = median(score0, na.rm=TRUE))
    
    meds2  <- meds %>%
    group_by(EventType) %>%
    dplyr::summarize(M2 = mean(M, na.rm=TRUE))
    
    dti$score1 <- NA
    for(ri in 1:max(dt$Experiment)){
      dti[dti$Experiment == ri, "score1"] <- myscale(x = dti[dti$Experiment == ri, "score0"],
                                                                   Syn.Avg.med = meds2$M2[meds2$EventType == "Synonymous"],
                                                                   StopGain.Avg.med = meds2$M2[meds2$EventType == "StopGain"],  
                                                                   Syn.med1 =  meds$M[meds$EventType == "Synonymous" & meds$Experiment == ri],
                                                                   StopGain.med1 = meds$M[meds$EventType == "StopGain" & meds$Experiment == ri])
    }
    dt.out <- rbind(dt.out, dti)
  }
  return(dt.out)
}
 

norm_x_exon <- function(dt){
  dt.out <- dt
  
  dti.temp <- dt %>% ## remove variants in splicing region
      filter(spliceR == FALSE) %>% filter(nchar(REF) == 1 & nchar(ALT) == 1) %>% ##filter(REF %in% c("A", "T", "C", "G") & ALT %in% c("A", "T", "C", "G") )%>%
    filter(inTarget == TRUE) 
  if(!is.null(params$vexclude)){
      dti.temp <- dti.temp %>% filter(!uPOS %in% Varaints_to_be_excluded_when_use_MAVE_internal_control_to_build_model$uPOS)  ## remove variants on chunling's list
  }
  meds <- dti.temp %>% group_by(Exon, EventType) %>%
    dplyr::summarize(M = median(score1, na.rm=TRUE))
  
  meds2  <- meds %>%
    group_by(EventType) %>%
    dplyr::summarize(M2 = mean(M, na.rm=TRUE))
  
  dt.out$score2 <- NA
  
  for(ei in unique(dt$Exon)){
    dt.out[dt.out$Exon == ei, "score2"] <- myscale(x = dt.out[dt.out$Exon == ei, "score1"],
                                                                   Syn.Avg.med = meds2$M2[meds2$EventType == "Synonymous"],
                                                                   StopGain.Avg.med = meds2$M2[meds2$EventType == "StopGain"],  
                                                                   Syn.med1 =  meds$M[meds$EventType == "Synonymous" & meds$Exon == ei  ],
                                                                   StopGain.med1 = meds$M[meds$EventType == "StopGain" & meds$Exon == ei])
  }
  
  return(dt.out)
}

## mydt has every variants
## mydt.snv has only SNVs in Target
## for summary numbers only, not used for later calculations
mydt.snv <- mydt %>% filter(nchar(REF) == 1 & nchar(ALT) == 1) %>% ##filter(REF %in% c("A", "T", "C", "G") & ALT %in% c("A", "T", "C", "G") ) %>% 
  filter(inTarget == TRUE)
```

## Results

In total, there are `r length(unique(dt.long$Exon))` Exons in this analysis. And EventType by spice region variants by Exon are 

```{r}

table( mydt.snv$Exon[!is.na(mydt.snv$score0)])
```

```{r}

table( mydt.snv$EventType[!is.na(mydt.snv$score0)], mydt.snv$Exon[!is.na(mydt.snv$score0)], mydt.snv$spliceR[!is.na(mydt.snv$score0)])
```

Since we have two/three or more replicates in each Exon, the data plotted below were doubled/tripled or multipled in number. 


Unique Variants are 

```{r}
## provide summary of number of variants.

dt.av <- mydt %>% dplyr::select(uPOS, Exon, REF, ALT, inTarget, functional.scoreA, functional.scoreB, functional.scoreC, functional.scoreD, spliceR, EventType) %>% unique()
dt <- dt.av %>% filter(nchar(REF) == 1 & nchar(ALT) == 1) %>%##filter(REF %in% c("A", "T", "C", "G") & ALT %in% c("A", "T", "C", "G") ) %>% 
  filter(inTarget == TRUE)
cat("score ", Score_method, "\n")
if(Score_method == "A"){
  print(table(dt$Exon[!is.na(dt$functional.scoreA)]))
  print(table( dt$EventType[!is.na(dt$functional.scoreA)], dt$Exon[!is.na(dt$functional.scoreA)], dt$spliceR[!is.na(dt$functional.scoreA)]))
}else if(Score_method == "B"){
  print(table(dt$Exon[!is.na(dt$functional.scoreB)]))
  print(table( dt$EventType[!is.na(dt$functional.scoreB)], dt$Exon[!is.na(dt$functional.scoreB)], dt$spliceR[!is.na(dt$functional.scoreB)]))
}else if(Score_method == "C"){
  print(table(dt$Exon[!is.na(dt$functional.scoreC)]))
  print(table( dt$EventType[!is.na(dt$functional.scoreC)], dt$Exon[!is.na(dt$functional.scoreC)], dt$spliceR[!is.na(dt$functional.scoreC)]))
}else{
  print(table(dt$Exon[!is.na(dt$functional.scoreD)]))
  print(table( dt$EventType[!is.na(dt$functional.scoreD)], dt$Exon[!is.na(dt$functional.scoreD)], dt$spliceR[!is.na(dt$functional.scoreD)]))
}
```

## 1. normalized Results for Ratio:


```{r, message=FALSE, warning=FALSE}

dt1 <- norm_win_exon(mydt)
dt2 <- norm_x_exon(dt1)


```


### Step1 normalization.

Plot shows SNV only

```{r, include=TRUE}
dt2.av <- dt2 ## dt2.av all variants, dt2 SNV only and in Target region

dt2 <- dt2.av %>% filter(nchar(REF) == 1 & nchar(ALT) == 1) %>% ##filter(REF %in% c("A", "T", "C", "G") & ALT %in% c("A", "T", "C", "G") ) %>% 
  filter(inTarget == TRUE)

```

```{r, include=TRUE, fig.width=3*min(length(exons), 3), fig.height=4*ceiling(length(exons)/3)}


etype8 <- unique(dt2$EventType)
if(length(etype8) >= 8){
  mycol8 <-  c(1, 5, 6, 2, 3, 4, 7, 8, 9)
}else{
  mycol8 <-  c(1, 5, 3, 4, 2, 6)

}
names(mycol8) <- etype8
mycol2.1 <- mycol8[etype8]

nrows <- max(ceiling(length(exons)/3), 1)
# debug only heeeeer
##nrows <- 1
par(mfrow=c(nrows, 3))
for(ei in unique(dt2$Exon)){
plot(dt2$score0[dt2$Exon == ei], dt2$score1[dt2$Exon == ei], pch = 20, col = mycol8[dt2$EventType[dt2$Exon == ei]], main = paste0(ei, " (N =", max(dt2$Experiment[dt2$Exon == ei & !is.na(dt2$score0)]), " , P =", length(unique(dt2$POS[dt2$Exon == ei])), ")"), 
     xlab = "score0", ylab = 'score1')
abline(a = 0, b = 1, col = "gray", lty = 2)
}
plot(0, 0, type = "n", xlab = '', ylab = '', axes = F)
legend(x = "bottom", 
       legend = etype8,
       pch =20,
       col =mycol2.1,
       xpd = TRUE, 
       cex = 1, ncol = 1) # Horizontal legend
```

### Step2 normalization.

Plot shows SNV only


```{r, include=TRUE, fig.width=3*min(length(exons)+1, 3), fig.height=4*ceiling(length(exons)/3)}

par(mfrow=c(nrows, 3))
for(ei in unique(dt$Exon)){
plot(dt2$score1[dt2$Exon == ei], dt2$score2[dt2$Exon == ei], pch = 20, col = mycol8[dt2$EventType[dt2$Exon == ei]], main = paste0(ei, " (N =", max(dt2$Experiment[dt2$Exon == ei & !is.na(dt2$score0)]), " , P =", length(unique(dt2$POS[dt2$Exon == ei])), ")"), 
     xlab = "score1", ylab = 'score2')
abline(a = 0, b = 1, col = "gray", lty = 2)
}

plot(0, 0, type = "n", xlab = '', ylab = '', axes = F)
legend(x = "bottom", 
       legend = etype8,
       pch =20,
       col =mycol2.1,
       xpd = TRUE, 
       cex = 1, ncol = 1) # Horizontal legend
```

### Put together 

Plot shows SNV only


```{r, include=TRUE, fig.width=12, fig.height=12}

par(mfrow=c(3,1))
plot(dt2$POS, dt2$score0, pch = 20, col = mycol8[dt2$EventType], main = "Raw", 
     xlab = "AA position", ylab = 'functional score', ylim = c(-7, 3))
abline(h = 0, col = "gray", lty = 2)
plot(dt2$POS, dt2$score1, pch = 20, col = mycol8[dt2$EventType], main = "Normalization within Exon", 
     xlab = "AA position", ylab = 'functional score', ylim = c(-7, 3))
abline(h = 0, col = "gray", lty = 2)
plot(dt2$POS, dt2$score2, pch = 20, col = mycol8[dt2$EventType], main = "Normalization across Exons", 
     xlab = "AA position", ylab = 'functional score', ylim = c(-7, 3))
abline(h = 0, col = "gray", lty = 2)
legend(x = "bottom", 
       legend = etype8,
       pch =20,
       col =mycol2.1,
       xpd = TRUE, 
       cex = 1, ncol = 3) # Horizontal legend

```


## 2. Score distribution and CI

### Norm check Individual scores ie pooled replicates

Plot shows SNV only

Including splicing region variants

```{r}

dt2plot <- dt2 %>% filter(EventType %in% c("StopGain", "Synonymous")) 
p <- dt2plot %>%
  ggplot( aes(x=score2, fill=EventType)) +
  geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity', binwidth = 0.20) +
  scale_fill_manual(values=c( "red", "#69b3a2")) +  ggtitle ( paste0(Ratio," individual score after normalization"))
ggplotly(p)


```

Plot shows SNV only

Excluding splicing region variants. Defined as SpliceR == TRUE only.

```{r}
dt2plot <- dt2 %>% filter(EventType %in% c("StopGain", "Synonymous")) %>% filter(spliceR== FALSE)
p <- dt2plot %>%
  ggplot( aes(x=score2, fill=EventType)) +
  geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity', binwidth = 0.2) +
  scale_fill_manual(values=c( "red", "#69b3a2")) +  ggtitle ( paste0(Ratio," individual score after normalization excluding splicing region"))
ggplotly(p)


```

Totally, there were `r nrow(dt2plot)` variants left from original `r nrow(dt2)` variants. 

```{r}

##write(length(which(dt2$EventType == "Synonymous")), stderr())

temp <- dt2$score2[dt2$EventType == "StopGain"]
shapiro.test(temp[sample(1:length(temp), min(4500, length(temp)))])

temp <- dt2$score2[dt2$EventType == "Synonymous"]
shapiro.test(temp[sample(1:length(temp), min(4500, length(temp)))])

```

### Norm check Average score ie single variant score

Plots show SNV only.

Including splice region variants.

```{r}
temp <- dt2.av %>% select(Exon, uPOS, POS, EventType, REF,     
ALT, AApos,AltAA, AltCodon, EventType, 
RefAA, RefCodon, SpliceAI_ALLELE, SpliceAI_DP_AG, SpliceAI_DP_AL, SpliceAI_DP_DG,  SpliceAI_DP_DL, SpliceAI_DS_AG,  SpliceAI_DS_AL, SpliceAI_DS_DG, SpliceAI_DS_DL, spliceR , Experiment, score2, inTarget) %>% unique() %>% mutate(Experiment = paste0("E", Experiment)) %>% tidyr::spread(key = Experiment, value = score2)


if(nmax == 6){
  D14vsLib_fs.2 <- temp %>%rowwise() %>% mutate(functional.score = mean(c(E1, E2, E3,E4, E5, E6), na.rm = T))
}else if(nmax == 5){
  D14vsLib_fs.2 <- temp %>%rowwise() %>% mutate(functional.score = mean(c(E1, E2, E3,E4,E5), na.rm = T))
}else if(nmax == 4){
  D14vsLib_fs.2 <- temp %>%rowwise() %>% mutate(functional.score = mean(c(E1, E2, E3,E4), na.rm = T))
}else if(nmax == 3){
  D14vsLib_fs.2 <- temp %>%rowwise() %>% mutate(functional.score = mean(c(E1, E2, E3), na.rm = T))
}else if(nmax == 2){
  D14vsLib_fs.2 <- temp %>%rowwise() %>% mutate(functional.score = mean(c(E1, E2), na.rm = T))
} else {
  D14vsLib_fs.2 <- temp %>%rowwise() %>% mutate(functional.score = E1)
}

temp1 <- dt2.av %>% select(uPOS, Experiment, score1) %>% mutate(Experiment = paste0("E", Experiment)) %>% tidyr::spread(key = Experiment, value = score1)
dt2.av.temp <- temp1
for (i in 1:nmax){
  Ei <- paste0("E", i)
  Ri_wie_norm <- paste0("R",i,"_wie_norm")
  dt2.av.temp <- dt2.av.temp %>% rename(!!sym(Ri_wie_norm) := !!sym(Ei))
}
temp1 <- dt2.av %>% select(uPOS, Experiment, score2) %>% mutate(Experiment = paste0("E", Experiment)) %>% tidyr::spread(key = Experiment, value = score2)
dt2.av.temp2 <- temp1
for (i in 1:nmax){
  Ei <- paste0("E", i)
  Ri_ce_norm <- paste0("R",i,"_ce_norm")
  dt2.av.temp2 <- dt2.av.temp2 %>% rename(!!sym(Ri_ce_norm) := !!sym(Ei))
}


dt2.av.temp <- merge(merge(dt0, dt2.av.temp), dt2.av.temp2) ## this is to be used to merge with D14vsLib_fs.2 later for output

# write(dim(D14vsLib_fs.2), stderr())
# write(length(unique(D14vsLib_fs.2$uPOS)), stderr())

D14vsLib_fs.2.av <- D14vsLib_fs.2
D14vsLib_fs.2 <- D14vsLib_fs.2.av %>% filter(nchar(REF) == 1 & nchar(ALT) == 1) %>% ##filter(REF %in% c("A", "T", "C", "G") & ALT %in% c("A", "T", "C", "G") ) %>% 
  filter(inTarget == TRUE)
dt2plot <- D14vsLib_fs.2 %>% filter(EventType %in% c("StopGain", "Synonymous")) 
p <- dt2plot %>% mutate(name = paste(Exon, uPOS)) %>% 
  ggplot( aes(x=functional.score, fill=EventType)) +
  geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity', binwidth = 0.1) +
  scale_fill_manual(values=c( "red", "#69b3a2")) +  ggtitle ( paste0("Functional.score ", Score_method, ifelse(Lab_curation == "No", " without ", " with "), "Lab Curation ", "extra filter applied_", extraFilter))
ggplotly(p)


```

Excluding splice region variants (spliceR == TRUE).

```{r}

dt2plot <- D14vsLib_fs.2%>% filter(EventType %in% c("StopGain", "Synonymous")) %>% filter(spliceR== FALSE)
p <- dt2plot %>%
  ggplot( aes(x=functional.score, fill=EventType)) +
  geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity', binwidth = 0.1) +
  scale_fill_manual(values=c( "red", "#69b3a2")) +  ggtitle ( paste0("Final.functional.score ", Score_method, ifelse(Lab_curation == "No", " w.o.", "with"), " Lab Curation", " extra filter applied_", extraFilter, " Excluding Splice region"))
ggplotly(p)


```

Functional scores By Exon, SNV only

```{r}
dt2plot <- D14vsLib_fs.2%>% filter(EventType %in% c("StopGain", "Synonymous")) %>% filter(spliceR== FALSE)
p <- dt2plot %>%
  ggplot( aes(x=functional.score, y = Exon, fill=EventType, text = uPOS)) +
  geom_point( color="#e9ecef", alpha=0.6, position = 'identity', binwidth = 0.1) +
  scale_fill_manual(values=c( "red", "#69b3a2")) +  ggtitle ( paste0("Final.functional.score ", Score_method, ifelse(Lab_curation == "No", " w.o.", "with"), " Lab Curation", " extra filter applied_", extraFilter, " Excluding Splice region"))
ggplotly(p)


```

```{r}
temp <- D14vsLib_fs.2
shapiro.test(temp$functional.score[temp$EventType == "StopGain"])
shapiro.test(temp$functional.score[temp$EventType == "Synonymous"])

```


### QQ plots 1. Individual scores

Plots show SNV only.

```{r, fig.width=8, fig.height=4}

par(mfrow=c(1,2))
qqnorm(dt2$score2[dt2$EventType == "StopGain"], main=paste0(Ratio, ' StopGain'))
qqline(dt2$score2[dt2$EventType == "StopGain"], col = "steelblue", lwd = 2)
qqnorm(dt2$score2[dt2$EventType == "Synonymous"], main=paste0(Ratio, ' Synonymous'))
qqline(dt2$score2[dt2$EventType == "Synonymous"], col = "steelblue", lwd = 2)


```

### QQ plots 2. Average score

Plots show SNV only

```{r, fig.width=8, fig.height=4}
temp <- D14vsLib_fs.2
par(mfrow=c(1,2))
qqnorm(temp$functional.score[temp$EventType == "StopGain"], main=paste0(Ratio, ' StopGain'))
qqline(temp$functional.score[temp$EventType == "StopGain"], col = "steelblue", lwd = 2)
qqnorm(temp$functional.score[temp$EventType == "Synonymous"], main=paste0(Ratio, ' Synonymous'))
qqline(temp$functional.score[temp$EventType == "Synonymous"], col = "steelblue", lwd = 2)

```

### Outlier defn 

Individual Score pooled are used. 

Focus on StopGain and Synonymous only.

```{r}

#method1 bootstrap

bootxP<-function(x,y, b,alpha){
# x is your data
# b is the number of bootstraps.
# alpha is your type I error


n <-2000
set.seed(123)
xp <- rep(0, b)
for(i in 1:b){
  x_boot<-c(sample(x,n,replace=TRUE), sample(y,n,replace=TRUE))
  xp[i] <- quantile(x_boot, alpha)
}

return(median(xp))
}

```



Bootstrap Method

1. Estimate of percentile of sample 

We sample large number (N) of samples and each time we get the xth percentile of the bootstrapped sample, estimate of the percentile will be the median of the N number of xth percentile. 


```{r}
x<- dt2$score2[dt2$EventType %in% c("StopGain")]
y<- dt2$score2[dt2$EventType %in% c("Synonymous")]

if(sum(is.na(x))>0) {
  x <- x[-which(is.na(x))]
}
if(sum(is.na(y))>0) {
  y <- y[-which(is.na(y))]
}

bl <- bu <- NA
if(params$Outlierremove != "No"){
  alpha <- ifelse(params$Outlierremove=="p1", 0.01, 
                  ifelse(params$Outlierremove == "p2", 0.02,
                         ifelse(params$Outlierremove == "p5", 0.05, NA)))
  bl <- bootxP(x, y,5000,alpha)
  bu <- bootxP(x, y,5000,1-alpha)
}

```

Plots show SNV only

Including splicing region variants.

```{r}

dt2plot <- dt2 %>% filter(EventType %in% c("Missense", "StopGain", "Synonymous"))
p <- dt2plot %>%
  ggplot( aes(x=score2, fill=EventType)) +
  geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity', binwidth = 0.1) +
  scale_fill_manual(values=c("blue", "red", "#69b3a2")) +  ggtitle ( paste0("score2", Score_method, ifelse(Lab_curation == "No", "without", "with"), "Lab Cureation"))
if(params$Outlierremove != "No"){
  p <- p+ geom_vline(xintercept=c(bl,bu), linetype="dashed", color = "red")  ## upper stopgain
}
ggplotly(p)


```

Excluding splicing region variants (spliceR == TRUE).

```{r}

dt2plot <- dt2 %>% filter(EventType %in% c("Missense", "StopGain", "Synonymous")) %>% filter(spliceR== FALSE)
p <- dt2plot %>%
  ggplot( aes(x=score2, fill=EventType)) +
  geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity', binwidth = 0.1) +
  scale_fill_manual(values=c("blue", "red", "#69b3a2")) +  ggtitle ( paste0("indiviual score after normalization ", Score_method, ifelse(Lab_curation == "No", "without", "with"), "Lab Cureation", "extra filter applied_", extraFilter))
if(params$Outlierremove != "No"){
  p <- p+ geom_vline(xintercept=c(bl,bu), linetype="dashed", color = "red")  ## upper stopgain
}

ggplotly(p)

```


`r ifelse(params$Outlierremove != "No", print(paste0(Ratio, " bootstrapped ", alpha*100, " and ", (1-alpha)*100, " percentile of StopGain & Synonymous are ", bl, " and ", bu)), "")`

```{r removeOutlier}

if(params$Outlierremove != "No"){
  dt.temp <- dt2.av %>% select(uPOS, Experiment, score2) %>% mutate(Experiment = paste0("S", Experiment)) %>% spread(key = Experiment, value = score2)
  dt0 <- merge(dt0, dt.temp)
  temp <- dt2.av %>% mutate(score2 = ifelse(score2 <bl |score2 >bu, NA, score2))
  temp <- temp %>% select(Exon, uPOS, POS, EventType, REF,     
                          ALT, AApos,AltAA, AltCodon, EventType, RefAA, RefCodon, SpliceAI_ALLELE, SpliceAI_DP_AG, SpliceAI_DP_AL, SpliceAI_DP_DG,  SpliceAI_DP_DL, SpliceAI_DS_AG,  SpliceAI_DS_AL, SpliceAI_DS_DG,
                          SpliceAI_DS_DL, spliceR , Experiment, score2) %>% mutate(Experiment = paste0("E", Experiment)) %>% spread(key = Experiment, value = score2)

  if(nmax == 6){
    D14vsLib_fs.2o <- temp %>%rowwise() %>% mutate(functional.score = mean(c(E1, E2, E3,E4, E5, E6), na.rm = T))
  }else if(nmax == 5){
    D14vsLib_fs.2o <- temp %>%rowwise() %>% mutate(functional.score = mean(c(E1, E2, E3,E4,E5), na.rm = T))
  }else if(nmax == 4){
    D14vsLib_fs.2o <- temp %>%rowwise() %>% mutate(functional.score = mean(c(E1, E2, E3,E4), na.rm = T))
  }else if(nmax == 3){
  D14vsLib_fs.2 <- temp %>%rowwise() %>% mutate(functional.score = mean(c(E1, E2, E3), na.rm = T))
}else if(nmax == 2){
  D14vsLib_fs.2o <- temp %>%rowwise() %>% mutate(functional.score = mean(c(E1, E2), na.rm = T))
} else {
  D14vsLib_fs.2o <- temp %>%rowwise() %>% mutate(functional.score = E1)
}
  D14vsLib_fs.2 <- D14vsLib_fs.2o
}


```

## 3. Group distribution

### Average

Plots show SNV only.

```{r}
dt2plot <- D14vsLib_fs.2 %>% filter(EventType %in% c("Missense", "StopGain", "Synonymous")) 

p <- dt2plot %>%
  ggplot( aes(x=functional.score, fill=EventType)) +
  geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity', binwidth = 0.10) +
  scale_fill_manual(values=c( "blue", "red","#69b3a2")) +  ggtitle ( paste0(Ratio," functional score"))
ggplotly(p)

dt2plot <-  D14vsLib_fs.2%>% filter(EventType %in% c("Missense", "StopGain", "Synonymous"))

ggplot(dt2plot, aes(x=EventType, y=functional.score)) + 
  geom_boxplot(outlier.colour="red", outlier.shape=8,
                outlier.size=4) +  ggtitle (Ratio) + geom_jitter(shape=16, position=position_jitter(0.2))

sg <- D14vsLib_fs.2$functional.score[D14vsLib_fs.2$EventType == "StopGain"]
sn <- D14vsLib_fs.2$functional.score[D14vsLib_fs.2$EventType == "Synonymous"]
ms <- D14vsLib_fs.2$functional.score[D14vsLib_fs.2$EventType == "Missense"]

wilcox.test(sg, sn, alternative = "l")
wilcox.test(ms, sn, alternative = "l")
wilcox.test(sg, ms, alternative = "l")

```

## 4. Classification perf 

This section we exclude the splice region variants (spliceR == TRUE) and variants on chunling's list

### 4.1 ROC average 

```{r rocfun}
## input score and type column name, casename, controlname 
## plot ROC, get auc, sens, spec
##
myroc2 <- function(data, typecolumn, casen, controln){
  data_long <- data %>% filter(spliceR == FALSE) %>% filter(nchar(REF) == 1 & nchar(ALT) == 1) %>% ##filter(REF %in% c("A", "T", "C", "G") & ALT %in% c("A", "T", "C", "G")) %>%
    filter( !!sym(typecolumn)%in% c(casen, controln)) %>% arrange(functional.score)
  ## plot
  par(mfrow = c(1, 1))
  roc_score = roc(controls = data_long$functional.score[which(data_long[,typecolumn] == controln)], cases = data_long$functional.score[which(data_long[,typecolumn] == casen)], direction = ">")  # AUC score
  plot(roc_score, main="ROC curve")
  text(0.2, 0.2, paste0("AUC: ", round(roc_score$auc, 3)))
  
  out <- data.frame(functional.score = roc_score$"threshold", sens =  roc_score$"sensitivities", specs = roc_score$"specificities")
  
  out$avg <- out$sens+out$specs
  
  ind <- which(out$avg == max(out$avg))
  if(length(ind) == 1){
    cutoff0A <- unique(out$functional.score[ind])
    sen0A <- out$sens[ind]
    spec0A <- out$specs[ind]
  
  } else{
    cutoff0A <- unique(out$functional.score[ind[1]])
    sen0A <- out$sens[ind[1]]
    spec0A <- out$specs[ind[1]]
  }
  
  auc0A <- roc_score$auc
  
  data_long$pred.class <- ifelse(data_long$functional.score < cutoff0A, casen, controln)
  cat("\n")
  
  x <- pull(data_long, typecolumn); y = data_long$pred.class
  x <- factor(x, levels = c(casen, controln)); y <- factor(y, levels = c(casen, controln))
  tb0A <- table(truth = x, pred = y)
  
  ## return the metrics
  return(list(n = nrow(data_long), cutoff = cutoff0A, auc = as.numeric(auc0A), sen = sen0A, spec = spec0A, tb = tb0A))
}

```

#### Data Driven

```{r ROC1, include=TRUE, fig.width=5, fig.height=5, message=FALSE, warning=FALSE}


temp <- D14vsLib_fs.2 %>% filter(EventType %in% c("StopGain", "Synonymous")) %>% filter(spliceR== FALSE) %>% arrange(functional.score)
if(!is.null(params$vexclude)){
    temp <- temp %>%  filter(!uPOS %in% Varaints_to_be_excluded_when_use_MAVE_internal_control_to_build_model$uPOS)  ## remove variants on chunling's list
}
ROC0 <- myroc2(temp,  typecolumn = "EventType", casen ="StopGain" , controln = "Synonymous")

```

Classification cutoff `r ROC0$cutoff`, sensitivity `r ROC0$sen`, specificity `r ROC0$spec`. 

`r print(ROC0$tb)`

#### Clinvar data

clinvar file is provided as `r params$clinvarfile`\

```{r}
if(!is.null(params$clinvarfile)){
  if (grepl("\\.txt$", params$clinvarfile)) {
    # Read the file using read.table
    DBD_missense_classified_in_ClinVar <- read.table(file = params$clinvarfile, header = TRUE, sep = '\t')
  } else {
    DBD_missense_classified_in_ClinVar <- read_excel(params$clinvarfile)
  }
  
  DBD_missense_classified_in_ClinVar <- DBD_missense_classified_in_ClinVar %>% unique()
  
  
  DBD_missense_classified_in_ClinVar <- DBD_missense_classified_in_ClinVar %>% mutate(uPOS = paste( GRCh38Location-1, Ref, Alt, sep = "_"))
  
  if(nrow(DBD_missense_classified_in_ClinVar) != length(unique(DBD_missense_classified_in_ClinVar$uPOS))) stop("error DBD_missense_classified_in_ClinVar has duplicate uPOS")
  
  if(length(grep("Exon", colnames(DBD_missense_classified_in_ClinVar)))>0){
    DBD_missense_classified_in_ClinVar <- DBD_missense_classified_in_ClinVar %>% dplyr::select(-Exon)
  }
  
  DBD_missense_classified_in_ClinVar$type <- ifelse(DBD_missense_classified_in_ClinVar$classification %in% c("B", "LB", "B (Single)", "B/LB", "LB (Single)"), "Benign", "Pathogenic")  
  
  D14vsLib_fs.2.av <- merge(D14vsLib_fs.2.av, DBD_missense_classified_in_ClinVar, by.x = "uPOS", by.y = "uPOS", all.x = T)

  D14vsLib_fs.2 <- D14vsLib_fs.2.av %>% filter(nchar(REF) == 1 & nchar(ALT) == 1) %>%##filter(REF %in% c("A", "T", "C", "G") & ALT %in% c("A", "T", "C", "G") )%>% 
    filter(inTarget == TRUE)
} else{
  D14vsLib_fs.2 <- D14vsLib_fs.2.av %>% filter(nchar(REF) == 1 & nchar(ALT) == 1) %>%##filter(REF %in% c("A", "T", "C", "G") & ALT %in% c("A", "T", "C", "G") )%>%
    filter(inTarget == TRUE)
}
```

```{r ROC1.c, include=TRUE, fig.width=5, fig.height=5, message=FALSE, warning=FALSE}

if(!is.null(params$clinvarfile)){
  temp <- D14vsLib_fs.2 %>% filter(type %in% c("Benign", "Pathogenic")) %>% filter(spliceR== FALSE) %>% arrange(functional.score)
  if(!is.null(params$vexclude)){
      temp <- temp %>%  filter(!uPOS %in% Varaints_to_be_excluded_when_use_MAVE_internal_control_to_build_model$uPOS)  ## remove variants on chunling's list
  }
  ROC0.c <- myroc2(temp,  typecolumn = "type", casen ="Pathogenic" , controln = "Benign")
}
```


`r ifelse(!is.null(params$clinvarfile), paste0("Classification cutoff ", ROC0.c$cutoff, ", sensitivity ", ROC0.c$sen, " specificity ", ROC0.c$spec), "")`

```{r}
if(!is.null(params$clinvarfile)) {
  print(ROC0.c$tb)
}
```

`r ifelse(!is.null(params$clinvarfile), "- apply data binary cutoff to clinvar data", "")`

```{r ROC1.dc, include=TRUE, fig.width=5, fig.height=5, message=FALSE, warning=FALSE}

if(!is.null(params$clinvarfile)) {
  temp <- D14vsLib_fs.2 %>% filter(type %in% c("Benign", "Pathogenic")) %>% filter(spliceR== FALSE) %>% arrange(functional.score) 
  
  temp$pred.class <- ifelse(temp$functional.score < ROC0$cutoff, "Pathogenic", "Benign")
  
  temp$pred.class <- factor(temp$pred.class, levels = c("Pathogenic", "Benign") )
  temp$type <- factor(temp$type, levels = c("Pathogenic", "Benign") )
  
  tb0.dc <- table(truth = temp$type, pred = temp$pred.class)
  tb0.dc
}

```


```{r, results = 'asis', fig.width=5, fig.height=5, message=FALSE, warning=FALSE}
if(!is.null(params$HDRfile)){
  if(nrow(HDR) != length(unique(HDR$uPOS))) stop("error HDR has duplicate uPOS")
  D14vsLib_fs.2.av <- merge(D14vsLib_fs.2.av, HDR, by.x = "uPOS", by.y = "uPOS", all.x = T)
  D14vsLib_fs.2 <- D14vsLib_fs.2.av %>% filter(nchar(REF) == 1 & nchar(ALT) == 1) %>%## filter(REF %in% c("A", "T", "C", "G") & ALT %in% c("A", "T", "C", "G") )%>%
    filter(inTarget == TRUE)
  
  cat("#### HDR data", "\n")
  cat("\n")
  temp <- D14vsLib_fs.2 %>% filter(htype %in% c("Abnormal", "Normal")) %>% filter(spliceR== FALSE) %>% arrange(functional.score)
  if(!is.null(params$vexclude)){
      temp <- temp %>%  filter(!uPOS %in% Varaints_to_be_excluded_when_use_MAVE_internal_control_to_build_model$uPOS)  ## remove variants on chunling's list
  }
  ROC0.h <- myroc2(temp,  typecolumn = "htype", casen ="Abnormal" , controln = "Normal")
  N_HDR <- nrow(temp)
  cat("\n Classification cutoff",  ROC0.h$cutoff, " sensitivity ", ROC0.h$sen, " specificity ",  ROC0.h$spec, ". \n\n" )
}
```

```{r}

if(!is.null(params$HDRfile)){
  print(ROC0.h$tb)
}
```

`r ifelse(!is.null(params$HDRfile), "-- apply data binary cutoff to HDR data ", "")`

```{r}
if(!is.null(params$HDRfile)){
  temp <- D14vsLib_fs.2 %>% filter(htype %in% c("Abnormal", "Normal")) %>% filter(spliceR== FALSE) %>% arrange(functional.score) 

  temp$pred.class <- ifelse(temp$functional.score < ROC0$cutoff, "Pathogenic", "Benign")

  temp$pred.class <- factor(temp$pred.class, levels = c("Pathogenic", "Benign") )
  temp$htype <- factor(temp$htype, levels = c("Abnormal", "Normal") )
  
  tb0.dh <- table(truth = temp$htype, pred = temp$pred.class)
  cat("\n")
  print(tb0.dh)
}
```

### 4.2 Mixture Model 

Findlay paper "fitting a two-component Gaussian mixture model in which the parameters of the ‘non-functional’ distribution were based on all nonsense SNVs and the ‘functional’ distribution based on synonymous SNVs not depleted in RNA (Extended Data Fig. 7). We then used this model to estimate the probability of each SNV’s score being drawn from the non-functional distribution (Pnf). SNVs with Pnf < 0.01 were categorized as functional (72.5%); SNVs with Pnf > 0.99 were categorized as non-functional (21.1%); and SNVs with 0.01 < Pnf < 0.99 (6.4%) were categorized as intermediate."



```{r MixModelfun}
mmfun2 <- function(data, typecolumn, casen, controln, lambda = "data", MM = "normal"){
  ## plot mixture model 
  ## Get thresholds and return posterior prob, 
  
  f1A <- f2A <- f3A <- f4A <- f5A <- f6A  <- NA
  
  data0 <- data %>% filter(!is.na(functional.score)) 
  # write(data0$uPOS[duplicated(data0$uPOS)], stderr())
  # write(data$uPOS[duplicated(data$uPOS)], stderr())
  #   
  
  temp <- data0 %>% filter(!!sym(typecolumn) %in% c(casen, controln)) %>% filter(nchar(REF) == 1 & nchar(ALT) == 1) %>% ##filter(REF %in% c("A", "T", "C", "G") & ALT %in% c("A", "T", "C", "G")) %>% 
    filter(spliceR == FALSE) %>% filter(inTarget == TRUE)

  if(!is.null(params$vexclude)){
    temp <- temp %>%  filter(!uPOS %in% Varaints_to_be_excluded_when_use_MAVE_internal_control_to_build_model$uPOS)  ## remove variants on chunling's list
  }
  
  myind <- match(temp$uPOS, data0$uPOS)
  
  m1 <- mean(temp$functional.score[which(pull(temp, typecolumn) == controln)], na.rm = TRUE)
  tmp <- temp$functional.score[which(pull(temp, typecolumn) == controln)]
  tmp2 <- tmp ## tmp[which(tmp > median(tmp)-sd(tmp) & tmp < median(tmp)+sd(tmp))]
  
  v1 <- sd(tmp2, na.rm = TRUE)
  
  m2 <- mean(temp$functional.score[which(pull(temp, typecolumn) == casen)])
  v2 <- sd(temp$functional.score[which(pull(temp, typecolumn) == casen)])
  l1 <- sum(pull(temp, typecolumn) == casen)/sum(pull(temp, typecolumn) == controln)
  
  if(params$MM == "Robust"){
    # robust calculation
    m2 = mean(temp$functional.score[which(pull(temp, typecolumn) == casen)], trim = 0.05)
    m1 = mean(temp$functional.score[which(pull(temp, typecolumn) == controln)], trim = 0.05)
    v2 = IQR(temp$functional.score[which(pull(temp, typecolumn) == casen)]) / 1.349
    v1 = IQR(temp$functional.score[which(pull(temp, typecolumn) == controln)]) / 1.349
  }
  error.occured <- FALSE
  tryCatch( {
    x <- normalmixEM(temp$functional.score, 
                     lambda = l1, k = 2,
                     mean.constr = c(m2, m1),
                     sd.constr = c(v2, v1),
                     mu = c(m2, m1), 
                     sigma = c(v2, v1))
    m1 = list(mu = x$mu, sd = x$sigma)
    
    prior = ifelse(lambda == "0.1", 0.1, ifelse(lambda == "0.2", 0.2, x$lambda[1]))
    print("prior used")
    print(prior)
    
    p1 = posteriorProbabilityFromGaussianModels(data0$functional.score, rep(0, length(data0$functional.score)), m1, prior = prior)
    
    # p1 has all prob and LR for non NA functional scores data0, m1 come from SNV, inTarget, non Splice region and subset to two levels only (functional and non functional)
    data0$posteriorP <- unlist(p1["posteriorP"])
    data0$likelihoodRatio  <- unlist(p1["likelihoodRatio"])
    
    temp2 <- left_join(data[,c("uPOS", "functional.score")], data0[, c("uPOS", "posteriorP", "likelihoodRatio")])
    data <- temp2 
    
    print("difference from data posterior < 1e-08: ")
    print(max(abs(unlist(p1['posteriorP'])[myind] -   x$posterior[,1])) < 1e-08)
    
    par(mfrow = c(1, 1))
    plot(temp$functional.score, unlist(p1['posteriorP'])[myind], col = ifelse(pull(temp, typecolumn) == casen, 2, 1))
    f1A <- temp$functional.score[unlist(p1['posteriorP'])[myind]>0.01][order(temp$functional.score[unlist(p1['posteriorP'])[myind]>0.01], decreasing = T)][1]
    f2A <- temp$functional.score[unlist(p1['posteriorP'])[myind]>0.99][order(temp$functional.score[unlist(p1['posteriorP'])[myind]>0.99], decreasing = T)][1]
    
    f3A <- temp$functional.score[unlist(p1['posteriorP'])[myind]>0.02][order(temp$functional.score[unlist(p1['posteriorP'])[myind]>0.02], decreasing = T)][1]
    f4A <- temp$functional.score[unlist(p1['posteriorP'])[myind]>0.98][order(temp$functional.score[unlist(p1['posteriorP'])[myind]>0.98], decreasing = T)][1]
    
    f5A <- temp$functional.score[unlist(p1['posteriorP'])[myind]>0.05][order(temp$functional.score[unlist(p1['posteriorP'])[myind]>0.05], decreasing = T)][1]
    f6A <- temp$functional.score[unlist(p1['posteriorP'])[myind]>0.95][order(temp$functional.score[unlist(p1['posteriorP'])[myind]>0.95], decreasing = T)][1]

    
    abline(v = f1A)
    abline(v = f2A)
    abline(h = 0.01)
    abline(h = 0.99)
    
    abline(v = f3A, lty = 2)
    abline(v = f4A, lty = 2)
    abline(h = 0.02, lty = 2)
    abline(h = 0.98, lty = 2)

    abline(v = f5A, lty = 3)
    abline(v = f6A, lty = 3)
    abline(h = 0.05, lty = 3)
    abline(h = 0.95, lty = 3)
    
    par(mfrow = c(1, 1))
    plot(x, which=2, xlab2= "", yaxt= "n", xlim = c(range(x$x)[1]-2, range(x$x)[2]+2), main2 ="Two components prob(nonfunctional) with cutoffs", lwd2=0.8, border = "azure3")
    lines(density(x$x), lty=2, lwd=0.8)
    
    abline(v = f1A)
    abline(v = f2A)
    abline(v = f3A, lty = 2)
    abline(v = f4A, lty = 2)
    
    abline(v = f5A, lty = 3)
    abline(v = f6A, lty = 3)
  }
  , error = function(e) {error.occured <<- TRUE})
  print(error.occured)
  
  return(list(error.occured = error.occured, f1 = as.numeric(f1A), f99 = as.numeric(f2A), f2=as.numeric(f3A), f98 = as.numeric(f4A), f5=as.numeric(f5A), f95 = as.numeric(f6A), data = data[, c("uPOS", "posteriorP", "likelihoodRatio")]))
}

```

#### Data driven

Based on distribution from StopGain and synonymous non splicing region (spliceR == FALSE), we also build mixture model.

```{r MixModel, include=TRUE, fig.width=9, fig.height=5, message=FALSE, warning=FALSE}
library(mixtools)

temp <- D14vsLib_fs.2 %>% filter(!is.na(functional.score)) %>% filter(EventType %in% c("StopGain", "Synonymous")) %>% 
  arrange(functional.score) %>% filter(spliceR== FALSE) %>% filter(inTarget == TRUE)
if(!is.null(params$vexclude)){
    temp <- temp %>%  filter(!uPOS %in% Varaints_to_be_excluded_when_use_MAVE_internal_control_to_build_model$uPOS)  ## remove variants on chunling's list
}

p <- temp %>%
  ggplot( aes(x=functional.score, fill=EventType)) +
  geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity', binwidth = 0.1) +
  scale_fill_manual(values=c("red",  "#69b3a2")) +  ggtitle ( "functional.score for MM building")
ggplotly(p)

D14vsLib_fs.2.av <- D14vsLib_fs.2.av %>% arrange(uPOS)
# write(D14vsLib_fs.2.av$uPOS[duplicated(D14vsLib_fs.2.av$uPOS)], stderr())
# write(paste0(params$Output_dir, Gene, "_", Ratio, "_", params$vset,"_", params$loessSubset, "_", params$EventCount_Filter, "_", params$Score_method, "_", (paste0(ifelse(Lab_curation == "No", "without", "with"), "_LabCureation_")), "E", gsub(", ", "", Exons), "_MM_", params$MM, "_Outlierremove_",  params$Outlierremove, "_extrafilter_", extraFilter, "_metrics.csv"), stderr()
# )
mm0 <- mmfun2(data = D14vsLib_fs.2.av, typecolumn = "EventType", casen = "StopGain", controln = "Synonymous", MM = params$MM, lambda = "data")
##write(dim(mm0$data$posteriorP), stderr())
##write(dim(D14vsLib_fs.2.av), stderr())
D14vsLib_fs.2.av$posteriorP <- mm0$data$posteriorP # <- D14vsLib_fs.2.av %>% mutate(posteriorP= mm0$data$posteriorP) 
```

`r paste0("There is",  ifelse(mm0$error.occured, " AN ", " NO "), "Error Ocurred")`.

`r mm0$f99` was the functional score corresponding to the probability of non functional at 99% and `r mm0$f1` was the functional score corresponding to the probability of non functional at 1%.

`r mm0$f98` was the functional score corresponding to the probability of non functional at 98% and `r mm0$f2` was the functional score corresponding to the probability of non functional at 2%.

`r mm0$f95` was the functional score corresponding to the probability of non functional at 95% and `r mm0$f5` was the functional score corresponding to the probability of non functional at 5%.


If we use these two thresholds to classify the variants of synonymous, stopgain and missense. 

First, 1-99% cutoff
Second, 2-98% cutoff
Third, 5-95% cutoff

Note: Splice regions are included in Missense, not stopgain nor synonymous
variants on chunling's list are included.

```{r MixModel2, include=TRUE, fig.width=9, fig.height=5, message=FALSE, warning=FALSE}
temp <- D14vsLib_fs.2 %>% filter(EventType %in% c("Synonymous", "StopGain")) %>% filter(spliceR == FALSE) 

temp2 <- D14vsLib_fs.2 %>% filter(EventType %in% c("Missense")) 

temp <- rbind(temp, temp2)


temp$ctype <- ifelse(temp$functional.score>mm0$f1, "neutural", 
                     ifelse(temp$functional.score>mm0$f99, "intermediate", "pathogenic"))


temp$ctype2 <- ifelse(temp$functional.score>mm0$f2, "neutural", 
                     ifelse(temp$functional.score>mm0$f98, "intermediate", "pathogenic"))

temp$ctype3 <- ifelse(temp$functional.score>mm0$f5, "neutural", 
                     ifelse(temp$functional.score>mm0$f95, "intermediate", "pathogenic"))


temp$ctype <- factor(temp$ctype, levels = c("pathogenic", "intermediate", "neutural"))
temp$ctype2 <- factor(temp$ctype2, levels = c("pathogenic", "intermediate", "neutural"))
temp$ctype3 <- factor(temp$ctype3, levels = c("pathogenic", "intermediate", "neutural"))

temp$EventType <-factor(temp$EventType, levels = c("StopGain", "Missense", "Synonymous"))

tb <- table(truth = temp$EventType,pred = temp$ctype)
print(tb)

tb2 <- table(truth = temp$EventType,pred = temp$ctype2)
print(tb2)
tb3 <- table(truth = temp$EventType,pred = temp$ctype3)
print(tb3)
```


` r ifelse(!is.null(params$clinvarfile), "- apply cutoffs to clinvar data:", "")` 


```{r MixModel2.dc, include=TRUE, fig.width=9, fig.height=5, message=FALSE, warning=FALSE}

if(!is.null(params$clinvarfile)) {
  temp <- D14vsLib_fs.2 %>% filter(type %in% c("Benign", "Pathogenic")) %>% filter(spliceR == FALSE) 
  
  temp$ctype <- ifelse(temp$functional.score>mm0$f1, "neutural", 
                       ifelse(temp$functional.score>mm0$f99, "intermediate", "pathogenic"))
  
  
  temp$ctype2 <- ifelse(temp$functional.score>mm0$f2, "neutural", 
                       ifelse(temp$functional.score>mm0$f98, "intermediate", "pathogenic"))
  
  temp$ctype3 <- ifelse(temp$functional.score>mm0$f5, "neutural", 
                       ifelse(temp$functional.score>mm0$f95, "intermediate", "pathogenic"))
  
  
  temp$ctype <- factor(temp$ctype, levels = c("pathogenic", "intermediate", "neutural"))
  temp$ctype2 <- factor(temp$ctype2, levels = c("pathogenic", "intermediate", "neutural"))
  temp$ctype3 <- factor(temp$ctype3, levels = c("pathogenic", "intermediate", "neutural"))
  
  tb <- table( truth = temp$type, pred=temp$ctype)
  print(tb)
  tb2 <- table( truth = temp$type, pred = temp$ctype2)
  print(tb2)
  tb3 <- table( truth = temp$type, pred = temp$ctype3)
  print(tb3)
}
```


#### Data driven with prior 0.1

Based on distribution from StopGain and synonymous non splicing region (spliceR == FALSE), we also build mixture model.

```{r MixModel.p01, include=TRUE, fig.width=9, fig.height=5, message=FALSE, warning=FALSE}

mm0.p01 <- mmfun2(data = D14vsLib_fs.2.av, typecolumn = "EventType", casen = "StopGain", controln = "Synonymous", MM = params$MM, lambda = "0.1")
D14vsLib_fs.2.av$posteriorP_p01= mm0.p01$data$posteriorP #<- D14vsLib_fs.2.av %>% mutate(posteriorP_p01= mm0.p01$data$posteriorP)
```

`r paste0("There is",  ifelse(mm0.p01$error.occured, " AN ", " NO "), "Error Ocurred")`.

`r mm0.p01$f99` was the functional score corresponding to the probability of non functional at 99% and `r mm0.p01$f1` was the functional score corresponding to the probability of non functional at 1%.

`r mm0.p01$f98` was the functional score corresponding to the probability of non functional at 98% and `r mm0.p01$f2` was the functional score corresponding to the probability of non functional at 2%.

`r mm0.p01$f95` was the functional score corresponding to the probability of non functional at 95% and `r mm0.p01$f5` was the functional score corresponding to the probability of non functional at 5%.


If we use these two thresholds to classify the variants of synonymous, stopgain and missense. 

First, 1-99% cutoff
Second, 2-98% cutoff
Third, 5-95% cutoff

Note: Splice regions are included in Missense, not stopgain nor synonymous
variants on chunling's list are included.

```{r MixModel2.p01, include=TRUE, fig.width=9, fig.height=5, message=FALSE, warning=FALSE}
temp <- D14vsLib_fs.2 %>% filter(EventType %in% c("Synonymous", "StopGain")) %>% filter(spliceR == FALSE) 

temp2 <- D14vsLib_fs.2 %>% filter(EventType %in% c("Missense")) 

temp <- rbind(temp, temp2)


temp$ctype <- ifelse(temp$functional.score>mm0.p01$f1, "neutural", 
                     ifelse(temp$functional.score>mm0.p01$f99, "intermediate", "pathogenic"))


temp$ctype2 <- ifelse(temp$functional.score>mm0.p01$f2, "neutural", 
                     ifelse(temp$functional.score>mm0.p01$f98, "intermediate", "pathogenic"))

temp$ctype3 <- ifelse(temp$functional.score>mm0.p01$f5, "neutural", 
                     ifelse(temp$functional.score>mm0.p01$f95, "intermediate", "pathogenic"))


temp$ctype <- factor(temp$ctype, levels = c("pathogenic", "intermediate", "neutural"))
temp$ctype2 <- factor(temp$ctype2, levels = c("pathogenic", "intermediate", "neutural"))
temp$ctype3 <- factor(temp$ctype3, levels = c("pathogenic", "intermediate", "neutural"))

temp$EventType <-factor(temp$EventType, levels = c("StopGain", "Missense", "Synonymous"))

tb <- table(truth = temp$EventType,pred = temp$ctype)
print(tb)

tb2 <- table(truth = temp$EventType,pred = temp$ctype2)
print(tb2)
tb3 <- table(truth = temp$EventType,pred = temp$ctype3)
print(tb3)
```


`r cat(ifelse(!is.null(params$clinvarfile), "- apply cutoffs to clinvar data: ", ""))`


```{r MixModel2.dc.p01, include=TRUE, fig.width=9, fig.height=5, message=FALSE, warning=FALSE}

if(!is.null(params$clinvarfile)) {
  temp <- D14vsLib_fs.2 %>% filter(type %in% c("Benign", "Pathogenic")) %>% filter(spliceR == FALSE) 
  
  temp$ctype <- ifelse(temp$functional.score>mm0.p01$f1, "neutural", 
                       ifelse(temp$functional.score>mm0.p01$f99, "intermediate", "pathogenic"))
  
  
  temp$ctype2 <- ifelse(temp$functional.score>mm0.p01$f2, "neutural", 
                       ifelse(temp$functional.score>mm0.p01$f98, "intermediate", "pathogenic"))
  
  temp$ctype3 <- ifelse(temp$functional.score>mm0.p01$f5, "neutural", 
                       ifelse(temp$functional.score>mm0.p01$f95, "intermediate", "pathogenic"))
  
  
  temp$ctype <- factor(temp$ctype, levels = c("pathogenic", "intermediate", "neutural"))
  temp$ctype2 <- factor(temp$ctype2, levels = c("pathogenic", "intermediate", "neutural"))
  temp$ctype3 <- factor(temp$ctype3, levels = c("pathogenic", "intermediate", "neutural"))
  
  tb <- table( truth = temp$type, pred=temp$ctype)
  print(tb)
  tb2 <- table( truth = temp$type, pred = temp$ctype2)
  print(tb2)
  tb3 <- table( truth = temp$type, pred = temp$ctype3)
  print(tb3)
}

```


#### Data driven with prior 0.2

Based on distribution from StopGain and synonymous non splicing region (spliceR == FALSE), we also build mixture model.

```{r MixModel.p02, include=TRUE, fig.width=9, fig.height=5, message=FALSE, warning=FALSE}

mm0.p02 <- mmfun2(data = D14vsLib_fs.2.av, typecolumn = "EventType", casen = "StopGain", controln = "Synonymous", MM = params$MM, lambda = "0.2")
D14vsLib_fs.2.av$posteriorP_p02= mm0.p02$data$posteriorP
D14vsLib_fs.2.av$likelihoodRatio= mm0.p02$data$likelihoodRatio ##<- D14vsLib_fs.2.av %>% mutate(posteriorP_p02= mm0.p02$data$posteriorP) %>% mutate(likelihoodRatio= mm0.p02$data$likelihoodRatio)
```

`r paste0("There is",  ifelse(mm0.p02$error.occured, " AN ", " NO "), "Error Ocurred")`.

`r mm0.p02$f99` was the functional score corresponding to the probability of non functional at 99% and `r mm0.p02$f1` was the functional score corresponding to the probability of non functional at 1%.

`r mm0.p02$f98` was the functional score corresponding to the probability of non functional at 98% and `r mm0.p02$f2` was the functional score corresponding to the probability of non functional at 2%.

`r mm0.p02$f95` was the functional score corresponding to the probability of non functional at 95% and `r mm0.p02$f5` was the functional score corresponding to the probability of non functional at 5%.


If we use these two thresholds to classify the variants of synonymous, stopgain and missense. 

First, 1-99% cutoff
Second, 2-98% cutoff
Third, 5-95% cutoff

Note: Splice regions are included in Missense, not stopgain nor synonymous
variants on chunling's list are included.

```{r MixModel2.p02, include=TRUE, fig.width=9, fig.height=5, message=FALSE, warning=FALSE}
temp <- D14vsLib_fs.2 %>% filter(EventType %in% c("Synonymous", "StopGain")) %>% filter(spliceR == FALSE) 

temp2 <- D14vsLib_fs.2 %>% filter(EventType %in% c("Missense")) 

temp <- rbind(temp, temp2)


temp$ctype <- ifelse(temp$functional.score>mm0.p02$f1, "neutural", 
                     ifelse(temp$functional.score>mm0.p02$f99, "intermediate", "pathogenic"))


temp$ctype2 <- ifelse(temp$functional.score>mm0.p02$f2, "neutural", 
                     ifelse(temp$functional.score>mm0.p02$f98, "intermediate", "pathogenic"))

temp$ctype3 <- ifelse(temp$functional.score>mm0.p02$f5, "neutural", 
                     ifelse(temp$functional.score>mm0.p02$f95, "intermediate", "pathogenic"))


temp$ctype <- factor(temp$ctype, levels = c("pathogenic", "intermediate", "neutural"))
temp$ctype2 <- factor(temp$ctype2, levels = c("pathogenic", "intermediate", "neutural"))
temp$ctype3 <- factor(temp$ctype3, levels = c("pathogenic", "intermediate", "neutural"))

temp$EventType <-factor(temp$EventType, levels = c("StopGain", "Missense", "Synonymous"))

tb <- table(truth = temp$EventType,pred = temp$ctype)
print(tb)

tb2 <- table(truth = temp$EventType,pred = temp$ctype2)
print(tb2)
tb3 <- table(truth = temp$EventType,pred = temp$ctype3)
print(tb3)
```


`r cat(ifelse(!is.null(params$clinvarfile), "- apply cutoffs to clinvar data: ", ""))`


```{r MixModel2.dc.p02, include=TRUE, fig.width=9, fig.height=5, message=FALSE, warning=FALSE}

if(!is.null(params$clinvarfile)){
  temp <- D14vsLib_fs.2 %>% filter(type %in% c("Benign", "Pathogenic")) %>% filter(spliceR == FALSE) 
  
  temp$ctype <- ifelse(temp$functional.score>mm0.p02$f1, "neutural", 
                       ifelse(temp$functional.score>mm0.p02$f99, "intermediate", "pathogenic"))
  
  
  temp$ctype2 <- ifelse(temp$functional.score>mm0.p02$f2, "neutural", 
                       ifelse(temp$functional.score>mm0.p02$f98, "intermediate", "pathogenic"))
  
  temp$ctype3 <- ifelse(temp$functional.score>mm0.p02$f5, "neutural", 
                       ifelse(temp$functional.score>mm0.p02$f95, "intermediate", "pathogenic"))
  
  
  temp$ctype <- factor(temp$ctype, levels = c("pathogenic", "intermediate", "neutural"))
  temp$ctype2 <- factor(temp$ctype2, levels = c("pathogenic", "intermediate", "neutural"))
  temp$ctype3 <- factor(temp$ctype3, levels = c("pathogenic", "intermediate", "neutural"))
  
  tb <- table( truth = temp$type, pred=temp$ctype)
  print(tb)
  tb2 <- table( truth = temp$type, pred = temp$ctype2)
  print(tb2)
  tb3 <- table( truth = temp$type, pred = temp$ctype3)
  print(tb3)
}

```

```{r}

if(!is.null(params$clinvarfile)){
  cat("\n")
  cat("#### ClinVar Data driven\n")
  
  cat("Based on distribution from Clinvar (pathogentic) and (benign), we also build mixture model.\n")
  
  cat("Exclude the spliceR and variants on chunling's list when build model.\n")
}
```

```{r}
if(!is.null(params$clinvarfile)){
  temp <- D14vsLib_fs.2 %>% filter(!is.na(functional.score)) %>% arrange(functional.score) %>% filter(type %in% c("Pathogenic", "Benign")) %>% filter(spliceR == FALSE) 
  if(!is.null(params$vexclude)){
      temp <- temp %>%  filter(!uPOS %in% Varaints_to_be_excluded_when_use_MAVE_internal_control_to_build_model$uPOS)  ## remove variants on chunling's list
  }
  
  temp$type <- factor(temp$type, levels = c("Pathogenic", "Benign"))
  
  cat('1. Clinvar variants in our Data\n')
  
  table( temp$EventType, temp$type, temp$Exon)
  
  cat('2. Use Clinvar variants for mixture model\n')
  
  table(temp$type)
}
```


```{r MixModel.C1, include=TRUE, fig.width=9, fig.height=5, message=FALSE, warning=FALSE}
if(!is.null(params$clinvarfile)){
  p <- temp %>%
    ggplot( aes(x=functional.score, fill=type)) +
    geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity', binwidth = 0.1) +
    scale_fill_manual(values=c(  "#69b3a2", "red")) +  ggtitle ( "functional.score by clinvar classification")
  ggplotly(p)
  
  
  
  mm0.c <- mmfun2(data = D14vsLib_fs.2.av, typecolumn = "type", casen = "Pathogenic", controln = "Benign", MM = params$MM, lambda = "data")
  D14vsLib_fs.2.av <- D14vsLib_fs.2.av %>% mutate(CposteriorP= mm0.c$data$posteriorP)
}
```


```{r}

if(!is.null(params$clinvarfile)){
   cat("\n")
  
  cat(paste0("There is",  ifelse(mm0.c$error.occured, " AN ", " NO "), "Error Ocurred\n"))
  

  cat(paste0(mm0.c$f99, " was the functional score corresponding to the probability of non functional at 99% and ", mm0.c$f1, " was the functional score corresponding to the probability of non functional at 1%.\n"))

  cat(paste0(mm0.c$f98, " was the functional score corresponding to the probability of non functional at 98% and ", mm0.c$f2, " was the functional score corresponding to the probability of non functional at 2%.\n"))

  cat(paste0(mm0.c$f95, " was the functional score corresponding to the probability of non functional at 95% and ", mm0.c$f5, " was the functional score corresponding to the probability of non functional at 5%.\n"))


  cat("If we use these two thresholds to classify the variants of synonymous, stopgain and missense. \n")
  
  cat("First, 1-99% cutoff\n")
  cat("Second, 2-98% cutoff\n")
  cat("Third, 5-95% cutoff\n")
  
  cat("Note: Splice regions are included in Missense, not stopgain nor synonymous variants on chunling's list are included.\n")

}

```

```{r MixModel2.C, include=TRUE, fig.width=9, fig.height=5, message=FALSE, warning=FALSE}
if(!is.null(params$clinvarfile)){
  if(!mm0.c$error.occured){
    temp <- D14vsLib_fs.2 %>% filter(EventType %in% c("Synonymous", "StopGain")) %>% filter(spliceR == FALSE) 
    
    temp2 <- D14vsLib_fs.2 %>% filter(EventType %in% c("Missense")) 
    
    temp <- rbind(temp, temp2)
    
    
    temp$ctype <- ifelse(temp$functional.score>mm0.c$f1, "neutural", 
                         ifelse(temp$functional.score>mm0.c$f99, "intermediate", "pathogenic"))
    
    
    temp$ctype2 <- ifelse(temp$functional.score>mm0.c$f2, "neutural", 
                         ifelse(temp$functional.score>mm0.c$f98, "intermediate", "pathogenic"))
    
    temp$ctype3 <- ifelse(temp$functional.score>mm0.c$f5, "neutural", 
                         ifelse(temp$functional.score>mm0.c$f95, "intermediate", "pathogenic"))
    
    
    temp$ctype <- factor(temp$ctype, levels = c("pathogenic", "intermediate", "neutural"))
    temp$ctype2 <- factor(temp$ctype2, levels = c("pathogenic", "intermediate", "neutural"))
    temp$ctype3 <- factor(temp$ctype3, levels = c("pathogenic", "intermediate", "neutural"))
    
    temp$EventType <-factor(temp$EventType, levels = c("StopGain", "Missense", "Synonymous"))
    
    tb <- table(truth = temp$EventType,pred = temp$ctype)
    print(tb)
    
    tb2 <- table(truth = temp$EventType,pred = temp$ctype2)
    print(tb2)
    tb3 <- table(truth = temp$EventType,pred = temp$ctype3)
    print(tb3)
  }
}
```

`r cat(ifelse(!is.null(params$clinvarfile),"- Apply cutoffs to clinvar data excluding spliceR\n", ""))`

```{r MixModel2.Cc, include=TRUE, fig.width=9, fig.height=5, message=FALSE, warning=FALSE}
if(!is.null(params$clinvarfile)){
  if(!mm0.c$error.occured){
    temp <- D14vsLib_fs.2 %>% filter(type %in% c("Benign", "Pathogenic")) %>% filter(spliceR == FALSE) 
    
    temp$ctype <- ifelse(temp$functional.score>mm0.c$f1, "neutural", 
                         ifelse(temp$functional.score>mm0.c$f99, "intermediate", "pathogenic"))
    
    
    temp$ctype2 <- ifelse(temp$functional.score>mm0.c$f2, "neutural", 
                         ifelse(temp$functional.score>mm0.c$f98, "intermediate", "pathogenic"))
    
    temp$ctype3 <- ifelse(temp$functional.score>mm0.c$f5, "neutural", 
                         ifelse(temp$functional.score>mm0.c$f95, "intermediate", "pathogenic"))
    
    
    temp$ctype <- factor(temp$ctype, levels = c("pathogenic", "intermediate", "neutural"))
    temp$ctype2 <- factor(temp$ctype2, levels = c("pathogenic", "intermediate", "neutural"))
    temp$ctype3 <- factor(temp$ctype3, levels = c("pathogenic", "intermediate", "neutural"))
    
    
    tb <- table(truth = temp$type, pred = temp$ctype)
    print(tb)
    
    tb2 <- table(truth = temp$type, pred = temp$ctype2)
    print(tb2)
    tb3 <- table( truth = temp$type, pred = temp$ctype3)
    print(tb3)
  }
}
```

```{r}
if(!is.null(params$clinvarfile)){
   cat("\n")
  
  cat("#### ClinVar Data driven with prior 0.1\n")
  
  cat("Based on distribution from Clinvar (pathogentic) and (benign), we also build mixture model.\n")
  
  cat("Exclude the spliceR and variants on chunling's list when build model.\n")
}
```


```{r}
if(!is.null(params$clinvarfile)){
  temp <- D14vsLib_fs.2 %>% filter(!is.na(functional.score)) %>% arrange(functional.score) %>% filter(type %in% c("Pathogenic", "Benign")) %>% filter(spliceR == FALSE) 
  if(!is.null(params$vexclude)){
      temp <- temp %>%  filter(!uPOS %in% Varaints_to_be_excluded_when_use_MAVE_internal_control_to_build_model$uPOS)  ## remove variants on chunling's list
  }
  
  temp$type <- factor(temp$type, levels = c("Pathogenic", "Benign"))
  
  cat("1. Clinvar variants in our Data\n")
  
  table(temp$EventType, temp$type, temp$Exon)
  
  cat("2. Use Clinvar variants for mixture model with prior 0.1\n")
  
  table(temp$type)
}

```



```{r MixModel.C1.p01, include=TRUE, fig.width=9, fig.height=5, message=FALSE, warning=FALSE}
if(!is.null(params$clinvarfile)){
mm0.c.p01 <- mmfun2(data = D14vsLib_fs.2.av, typecolumn = "type", casen = "Pathogenic", controln = "Benign", MM = params$MM, lambda = "0.1")
D14vsLib_fs.2.av <- D14vsLib_fs.2.av %>% mutate(CposteriorP_p01= mm0.c.p01$data$posteriorP)
}
```

```{r}
if(!is.null(params$clinvarfile)){
cat( paste0("There is",  ifelse(mm0.c.p01$error.occured, " AN ", " NO "), "Error Ocurred.\n"))

cat( paste0( mm0.c.p01$f99, " was the functional score corresponding to the probability of non functional at 99% and ", mm0.c.p01$f1, " was the functional score corresponding to the probability of non functional at 1%.\n"))

cat( paste0( mm0.c.p01$f98, " was the functional score corresponding to the probability of non functional at 98% and ", mm0.c.p01$f2, " was the functional score corresponding to the probability of non functional at 2%.\n"))

cat( paste0( mm0.c.p01$f95, " was the functional score corresponding to the probability of non functional at 95% and ", mm0.c.p01$f5, " was the functional score corresponding to the probability of non functional at 5%.\n"))


cat("If we use these two thresholds to classify the variants of synonymous, stopgain and missense. \n")

cat("First, 1-99% cutoff.\n")
cat("Second, 2-98% cutoff.\n")
cat("Third, 5-95% cutoff.\n")

cat("Note: Splice regions are included in Missense, not stopgain nor synonymous
variants on chunling's list are included.\n")

}

```

```{r MixModel2.C.p01, include=TRUE, fig.width=9, fig.height=5, message=FALSE, warning=FALSE}
if(!is.null(params$clinvarfile)){
  if(!mm0.c.p01$error.occured){
    temp <- D14vsLib_fs.2 %>% filter(EventType %in% c("Synonymous", "StopGain")) %>% filter(spliceR == FALSE) 
    
    temp2 <- D14vsLib_fs.2 %>% filter(EventType %in% c("Missense")) 
    
    temp <- rbind(temp, temp2)
    
    
    temp$ctype <- ifelse(temp$functional.score>mm0.c.p01$f1, "neutural", 
                         ifelse(temp$functional.score>mm0.c.p01$f99, "intermediate", "pathogenic"))
    
    
    temp$ctype2 <- ifelse(temp$functional.score>mm0.c.p01$f2, "neutural", 
                         ifelse(temp$functional.score>mm0.c.p01$f98, "intermediate", "pathogenic"))
    
    temp$ctype3 <- ifelse(temp$functional.score>mm0.c.p01$f5, "neutural", 
                         ifelse(temp$functional.score>mm0.c.p01$f95, "intermediate", "pathogenic"))
    
    
    temp$ctype <- factor(temp$ctype, levels = c("pathogenic", "intermediate", "neutural"))
    temp$ctype2 <- factor(temp$ctype2, levels = c("pathogenic", "intermediate", "neutural"))
    temp$ctype3 <- factor(temp$ctype3, levels = c("pathogenic", "intermediate", "neutural"))
    
    temp$EventType <-factor(temp$EventType, levels = c("StopGain", "Missense", "Synonymous"))
    
    tb <- table(truth = temp$EventType,pred = temp$ctype)
    print(tb)
    
    tb2 <- table(truth = temp$EventType,pred = temp$ctype2)
    print(tb2)
    tb3 <- table(truth = temp$EventType,pred = temp$ctype3)
    print(tb3)
  }
}
```

`r cat(ifelse(!is.null(params$clinvarfile), "- Apply cutoffs to clinvar data excluding spliceR", ""))`

```{r MixModel2.Cc.p01, include=TRUE, fig.width=9, fig.height=5, message=FALSE, warning=FALSE}
if(!is.null(params$clinvarfile)){
  
  if(!mm0.c.p01$error.occured){
    temp <- D14vsLib_fs.2 %>% filter(type %in% c("Benign", "Pathogenic")) %>% filter(spliceR == FALSE) 
    
    temp$ctype <- ifelse(temp$functional.score>mm0.c.p01$f1, "neutural", 
                         ifelse(temp$functional.score>mm0.c.p01$f99, "intermediate", "pathogenic"))
    
    
    temp$ctype2 <- ifelse(temp$functional.score>mm0.c.p01$f2, "neutural", 
                         ifelse(temp$functional.score>mm0.c.p01$f98, "intermediate", "pathogenic"))
    
    temp$ctype3 <- ifelse(temp$functional.score>mm0.c.p01$f5, "neutural", 
                         ifelse(temp$functional.score>mm0.c.p01$f95, "intermediate", "pathogenic"))
    
    
    temp$ctype <- factor(temp$ctype, levels = c("pathogenic", "intermediate", "neutural"))
    temp$ctype2 <- factor(temp$ctype2, levels = c("pathogenic", "intermediate", "neutural"))
    temp$ctype3 <- factor(temp$ctype3, levels = c("pathogenic", "intermediate", "neutural"))
    
    
    tb <- table(truth = temp$type, pred = temp$ctype)
    print(tb)
    
    tb2 <- table(truth = temp$type, pred = temp$ctype2)
    print(tb2)
    tb3 <- table( truth = temp$type, pred = temp$ctype3)
    print(tb3)
  }
}
```

```{r}

if(!is.null(params$clinvarfile)){
  
   cat("\n")
  
  cat("#### ClinVar Data driven with prior 0.2\n")

cat("Based on distribution from Clinvar (pathogentic) and (benign), we also build mixture model.\n")

cat("Exclude the spliceR and variants on chunling's list when build model.\n")
}

```

```{r}
if(!is.null(params$clinvarfile)){
  temp <- D14vsLib_fs.2 %>% filter(!is.na(functional.score)) %>% arrange(functional.score) %>% filter(type %in% c("Pathogenic","Benign")) %>% filter(spliceR == FALSE) 
  if(!is.null(params$vexclude)){
      temp <- temp %>%  filter(!uPOS %in% Varaints_to_be_excluded_when_use_MAVE_internal_control_to_build_model$uPOS)  ## remove variants on chunling's list
  }
  
  temp$type <- factor(temp$type, levels = c("Pathogenic", "Benign"))
  cat("1. Clinvar variants in our Data\n")
  
  table( temp$EventType, temp$type, temp$Exon)
  cat("2. Use Clinvar variants for mixture model\n")
  
  table(temp$type)
}

```

```{r MixModel.C1.p02, include=TRUE, fig.width=9, fig.height=5, message=FALSE, warning=FALSE}
if(!is.null(params$clinvarfile)){
  mm0.c.p02 <- mmfun2(data = D14vsLib_fs.2.av, typecolumn = "type", casen = "Pathogenic", controln = "Benign", MM = params$MM, lambda = "0.2")
  D14vsLib_fs.2.av <- D14vsLib_fs.2.av %>% mutate(CposteriorP_p02= mm0.c.p02$data$posteriorP) %>% mutate(ClikelihoodRatio= mm0.c.p02$data$likelihoodRatio)
}
```

```{r}

if(!is.null(params$clinvarfile)){
  cat("\n")
  cat(paste0("There is",  ifelse(mm0.c.p02$error.occured, " AN ", " NO "), "Error Ocurred.\n"))

  cat(paste0(mm0.c.p02$f99, " was the functional score corresponding to the probability of non functional at 99% and ", mm0.c.p02$f1, " was the functional score corresponding to the probability of non functional at 1%.\n"))

  cat(paste0( mm0.c.p02$f98, " was the functional score corresponding to the probability of non functional at 98% and ",  mm0.c.p02$f2, " was the functional score corresponding to the probability of non functional at 2%.\n"))

  cat(paste0(mm0.c.p02$f95, " was the functional score corresponding to the probability of non functional at 95% and ", mm0.c.p02$f5, " was the functional score corresponding to the probability of non functional at 5%.\n"))


  cat("If we use these two thresholds to classify the variants of synonymous, stopgain and missense. .\n")

  cat("First, 1-99% cutoff.\n")
  cat("Second, 2-98% cutoff.\n")
  cat("Third, 5-95% cutoff.\n")
  
  cat("Note: Splice regions are included in Missense, not stopgain nor synonymous variants on chunling's list are included..\n")
}

```


```{r MixModel2.C.p02, include=TRUE, fig.width=9, fig.height=5, message=FALSE, warning=FALSE}
if(!is.null(params$clinvarfile)){
  if(!mm0.c.p02$error.occured){
    temp <- D14vsLib_fs.2 %>% filter(EventType %in% c("Synonymous", "StopGain")) %>% filter(spliceR == FALSE) 
    
    temp2 <- D14vsLib_fs.2 %>% filter(EventType %in% c("Missense")) 
    
    temp <- rbind(temp, temp2)
    
    
    temp$ctype <- ifelse(temp$functional.score>mm0.c.p02$f1, "neutural", 
                         ifelse(temp$functional.score>mm0.c.p02$f99, "intermediate", "pathogenic"))
    
    
    temp$ctype2 <- ifelse(temp$functional.score>mm0.c.p02$f2, "neutural", 
                         ifelse(temp$functional.score>mm0.c.p02$f98, "intermediate", "pathogenic"))
    
    temp$ctype3 <- ifelse(temp$functional.score>mm0.c.p02$f5, "neutural", 
                         ifelse(temp$functional.score>mm0.c.p02$f95, "intermediate", "pathogenic"))
    
    
    temp$ctype <- factor(temp$ctype, levels = c("pathogenic", "intermediate", "neutural"))
    temp$ctype2 <- factor(temp$ctype2, levels = c("pathogenic", "intermediate", "neutural"))
    temp$ctype3 <- factor(temp$ctype3, levels = c("pathogenic", "intermediate", "neutural"))
    
    temp$EventType <-factor(temp$EventType, levels = c("StopGain", "Missense", "Synonymous"))
    
    tb <- table(truth = temp$EventType,pred = temp$ctype)
    print(tb)
    
    tb2 <- table(truth = temp$EventType,pred = temp$ctype2)
    print(tb2)
    tb3 <- table(truth = temp$EventType,pred = temp$ctype3)
    print(tb3)
  }
}
```

`r cat(ifelse(!is.null(params$clinvarfile), "- Apply cutoffs to clinvar data excluding spliceR", ""))`

```{r MixModel2.Cc.p02, include=TRUE, fig.width=9, fig.height=5, message=FALSE, warning=FALSE}
if(!is.null(params$clinvarfile)){
  if(!mm0.c.p02$error.occured){
    temp <- D14vsLib_fs.2 %>% filter(type %in% c("Benign", "Pathogenic")) %>% filter(spliceR == FALSE) 
    
    temp$ctype <- ifelse(temp$functional.score>mm0.c.p02$f1, "neutural", 
                         ifelse(temp$functional.score>mm0.c.p02$f99, "intermediate", "pathogenic"))
    
    
    temp$ctype2 <- ifelse(temp$functional.score>mm0.c.p02$f2, "neutural", 
                         ifelse(temp$functional.score>mm0.c.p02$f98, "intermediate", "pathogenic"))
    
    temp$ctype3 <- ifelse(temp$functional.score>mm0.c.p02$f5, "neutural", 
                         ifelse(temp$functional.score>mm0.c.p02$f95, "intermediate", "pathogenic"))
    
    
    temp$ctype <- factor(temp$ctype, levels = c("pathogenic", "intermediate", "neutural"))
    temp$ctype2 <- factor(temp$ctype2, levels = c("pathogenic", "intermediate", "neutural"))
    temp$ctype3 <- factor(temp$ctype3, levels = c("pathogenic", "intermediate", "neutural"))
    
    
    tb <- table(truth = temp$type, pred = temp$ctype)
    print(tb)
    
    tb2 <- table(truth = temp$type, pred = temp$ctype2)
    print(tb2)
    tb3 <- table( truth = temp$type, pred = temp$ctype3)
    print(tb3)
  }
}
```




```{r, results = 'asis',  fig.width=5, fig.height=5, message=FALSE, warning=FALSE}
if(!is.null(params$clinvarfile)){
  if(!is.null(params$HDRfile)){
   cat("\n")
    cat("#### HDR data", "\n")
    cat("\n")
    
    temp <- D14vsLib_fs.2 %>% filter(!is.na(functional.score)) %>% arrange(functional.score) %>% filter(htype %in% c("Abnormal", "Normal")) %>% filter(spliceR == FALSE) 
    if(!is.null(params$vexclude)){
        temp <- temp %>%  filter(!uPOS %in% Varaints_to_be_excluded_when_use_MAVE_internal_control_to_build_model$uPOS)  ## remove variants on chunling's list
    }
  }
}

```

`r ifelse(!is.null(params$HDRfile), " 1. HDR variants in our Data", "")`

```{r}
if(!is.null(params$HDRfile)){
  print(table( temp$EventType, temp$htype, temp$Exon))
}
```

`r ifelse(!is.null(params$HDRfile), " 2. Use HDR variants for mixture model" , "")`

```{r,  fig.width=9, fig.height=5, message=FALSE, warning=FALSE}
if(!is.null(params$HDRfile)){
  print(table(temp$htype))

  p <- temp %>%
    ggplot( aes(x=functional.score, fill=type)) +
    geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity', binwidth = 0.1) +
    scale_fill_manual(values=c(  "#69b3a2", "red")) +  ggtitle ( "functional.score by clinvar classification")
  ggplotly(p)

  mm0.h <- mmfun2(data = D14vsLib_fs.2.av, typecolumn = "htype", casen = "Abnormal", controln = "Normal", MM = params$MM, lambda = "data")
  D14vsLib_fs.2.av <- D14vsLib_fs.2.av %>% mutate(HposteriorP= mm0.h$data$posteriorP)

}
```

```{r, results = 'asis', message=FALSE, warning=FALSE}
if(!is.null(params$HDRfile)){

  
  cat(paste0("There is",  ifelse(mm0.h$error.occured, " AN ", " NO "), "Error Ocurred  \n"))
  cat(paste0(mm0.h$f99, " was the functional score corresponding to the probability of non functional at 99% and ",  mm0.h$f1, " was the functional score corresponding to the probability of non functional at 1%. \n"))

  cat(paste0(mm0.h$f98,"  was the functional score corresponding to the probability of non functional at 98% and ",  mm0.h$f2, " was the functional score corresponding to the probability of non functional at 2%. \n"))
  
  cat(paste0( mm0.h$f95, " was the functional score corresponding to the probability of non functional at 95% and ", mm0.h$f5," was the functional score corresponding to the probability of non functional at 5%. \n"))
  
  cat("\n If we use these two thresholds to classify the variants of synonymous, stopgain and missense. \n")

  cat("\n First, 1-99% cutoff . \n")
  cat("\n Second, 2-98% cutoff . \n")
  cat("\n Third, 5-95% cutoff . \n")
  
  cat("\n Note: Splice regions are included in Missense, not stopgain nor synonymous
  variants on chunling's list are included.. \n")

}
```

```{r, message=FALSE, warning=FALSE}
if(!is.null(params$HDRfile)){
  if(!mm0.h$error.occured){
    temp <- D14vsLib_fs.2 %>% filter(EventType %in% c("Synonymous", "StopGain")) %>% filter(spliceR == FALSE) 
    
    temp2 <- D14vsLib_fs.2 %>% filter(EventType %in% c("Missense")) 
    
    temp <- rbind(temp, temp2)
    
    
    temp$ctype <- ifelse(temp$functional.score>mm0.h$f1, "neutural", 
                         ifelse(temp$functional.score>mm0.h$f99, "intermediate", "pathogenic"))
    
    
    temp$ctype2 <- ifelse(temp$functional.score>mm0.h$f2, "neutural", 
                         ifelse(temp$functional.score>mm0.h$f98, "intermediate", "pathogenic"))
    
    temp$ctype3 <- ifelse(temp$functional.score>mm0.h$f5, "neutural", 
                         ifelse(temp$functional.score>mm0.h$f95, "intermediate", "pathogenic"))
    
    
    temp$ctype <- factor(temp$ctype, levels = c("pathogenic", "intermediate", "neutural"))
    temp$ctype2 <- factor(temp$ctype2, levels = c("pathogenic", "intermediate", "neutural"))
    temp$ctype3 <- factor(temp$ctype3, levels = c("pathogenic", "intermediate", "neutural"))
    
    temp$htype <- factor(temp$htype, levels = c("Abnormal", "Normal"))
    cat("\n")
    tb <- table(truth = temp$htype, pred = temp$ctype)
    print(tb)
    
    tb2 <- table(truth = temp$htype, pred = temp$ctype2)
    print(tb2)
    tb3 <- table(truth = temp$htype, pred = temp$ctype3)
    print(tb3)
  }

}

```

`r ifelse(!is.null(params$HDRfile), " 3. Use HDR variants for mixture model with prior 0.1" , "")`

```{r,  fig.width=9, fig.height=5, message=FALSE, warning=FALSE}
if(!is.null(params$HDRfile)){
  mm0.h.p01 <- mmfun2(data = D14vsLib_fs.2.av, typecolumn = "htype", casen = "Abnormal", controln = "Normal", MM = params$MM, lambda = "0.1")
  D14vsLib_fs.2.av <- D14vsLib_fs.2.av %>% mutate(HposteriorP_p01= mm0.h.p01$data$posteriorP)

}
```

```{r, results = 'asis', message=FALSE, warning=FALSE}
if(!is.null(params$HDRfile)){

  
  cat(paste0("There is",  ifelse(mm0.h.p01$error.occured, " AN ", " NO "), "Error Ocurred  \n"))
  cat(paste0(mm0.h.p01$f99, " was the functional score corresponding to the probability of non functional at 99% and ",  mm0.h.p01$f1, " was the functional score corresponding to the probability of non functional at 1%. \n"))

  cat(paste0(mm0.h.p01$f98,"  was the functional score corresponding to the probability of non functional at 98% and ",  mm0.h.p01$f2, " was the functional score corresponding to the probability of non functional at 2%. \n"))
  
  cat(paste0( mm0.h.p01$f95, " was the functional score corresponding to the probability of non functional at 95% and ", mm0.h.p01$f5," was the functional score corresponding to the probability of non functional at 5%. \n"))
  
  cat("\n If we use these two thresholds to classify the variants of synonymous, stopgain and missense. \n")

  cat("\n First, 1-99% cutoff . \n")
  cat("\n Second, 2-98% cutoff . \n")
  cat("\n Third, 5-95% cutoff . \n")
  
  cat("\n Note: Splice regions are included in Missense, not stopgain nor synonymous
  variants on chunling's list are included.. \n")

}
```

```{r, message=FALSE, warning=FALSE}
if(!is.null(params$HDRfile)){
  if(!mm0.h.p01$error.occured){
    temp <- D14vsLib_fs.2 %>% filter(EventType %in% c("Synonymous", "StopGain")) %>% filter(spliceR == FALSE) 
    
    temp2 <- D14vsLib_fs.2 %>% filter(EventType %in% c("Missense")) 
    
    temp <- rbind(temp, temp2)
    
    
    temp$ctype <- ifelse(temp$functional.score>mm0.h.p01$f1, "neutural", 
                         ifelse(temp$functional.score>mm0.h.p01$f99, "intermediate", "pathogenic"))
    
    
    temp$ctype2 <- ifelse(temp$functional.score>mm0.h.p01$f2, "neutural", 
                         ifelse(temp$functional.score>mm0.h.p01$f98, "intermediate", "pathogenic"))
    
    temp$ctype3 <- ifelse(temp$functional.score>mm0.h.p01$f5, "neutural", 
                         ifelse(temp$functional.score>mm0.h.p01$f95, "intermediate", "pathogenic"))
    
    
    temp$ctype <- factor(temp$ctype, levels = c("pathogenic", "intermediate", "neutural"))
    temp$ctype2 <- factor(temp$ctype2, levels = c("pathogenic", "intermediate", "neutural"))
    temp$ctype3 <- factor(temp$ctype3, levels = c("pathogenic", "intermediate", "neutural"))
    
    temp$htype <- factor(temp$htype, levels = c("Abnormal", "Normal"))
    cat("\n")
    tb <- table(truth = temp$htype, pred = temp$ctype)
    print(tb)
    
    tb2 <- table(truth = temp$htype, pred = temp$ctype2)
    print(tb2)
    tb3 <- table(truth = temp$htype, pred = temp$ctype3)
    print(tb3)
  }

}

```

`r ifelse(!is.null(params$HDRfile), " 4. Use HDR variants for mixture model with prior 0.2" , "")`

```{r,  fig.width=9, fig.height=5, message=FALSE, warning=FALSE}
if(!is.null(params$HDRfile)){

  mm0.h.p02 <- mmfun2(data = D14vsLib_fs.2.av, typecolumn = "htype", casen = "Abnormal", controln = "Normal", MM = params$MM, lambda = "0.1")
  D14vsLib_fs.2.av <- D14vsLib_fs.2.av %>% mutate(HposteriorP_p02= mm0.h.p02$data$posteriorP) %>% mutate(HlikelihoodRatio= mm0.h.p02$data$likelihoodRatio)

}
```

```{r, results = 'asis', message=FALSE, warning=FALSE}
if(!is.null(params$HDRfile)){

  
  cat(paste0("There is",  ifelse(mm0.h.p02$error.occured, " AN ", " NO "), "Error Ocurred  \n"))
  cat(paste0(mm0.h.p02$f99, " was the functional score corresponding to the probability of non functional at 99% and ",  mm0.h.p02$f1, " was the functional score corresponding to the probability of non functional at 1%. \n"))

  cat(paste0(mm0.h.p02$f98,"  was the functional score corresponding to the probability of non functional at 98% and ",  mm0.h.p02$f2, " was the functional score corresponding to the probability of non functional at 2%. \n"))
  
  cat(paste0( mm0.h.p02$f95, " was the functional score corresponding to the probability of non functional at 95% and ", mm0.h.p02$f5," was the functional score corresponding to the probability of non functional at 5%. \n"))
  
  cat("\n If we use these two thresholds to classify the variants of synonymous, stopgain and missense. \n")

  cat("\n First, 1-99% cutoff . \n")
  cat("\n Second, 2-98% cutoff . \n")
  cat("\n Third, 5-95% cutoff . \n")
  
  cat("\n Note: Splice regions are included in Missense, not stopgain nor synonymous
  variants on chunling's list are included.. \n")

}
```

```{r, message=FALSE, warning=FALSE}
if(!is.null(params$HDRfile)){
  if(!mm0.h.p02$error.occured){
    temp <- D14vsLib_fs.2 %>% filter(EventType %in% c("Synonymous", "StopGain")) %>% filter(spliceR == FALSE) 
    
    temp2 <- D14vsLib_fs.2 %>% filter(EventType %in% c("Missense")) 
    
    temp <- rbind(temp, temp2)
    
    
    temp$ctype <- ifelse(temp$functional.score>mm0.h.p02$f1, "neutural", 
                         ifelse(temp$functional.score>mm0.h.p02$f99, "intermediate", "pathogenic"))
    
    
    temp$ctype2 <- ifelse(temp$functional.score>mm0.h.p02$f2, "neutural", 
                         ifelse(temp$functional.score>mm0.h.p02$f98, "intermediate", "pathogenic"))
    
    temp$ctype3 <- ifelse(temp$functional.score>mm0.h.p02$f5, "neutural", 
                         ifelse(temp$functional.score>mm0.h.p02$f95, "intermediate", "pathogenic"))
    
    
    temp$ctype <- factor(temp$ctype, levels = c("pathogenic", "intermediate", "neutural"))
    temp$ctype2 <- factor(temp$ctype2, levels = c("pathogenic", "intermediate", "neutural"))
    temp$ctype3 <- factor(temp$ctype3, levels = c("pathogenic", "intermediate", "neutural"))
    
    temp$htype <- factor(temp$htype, levels = c("Abnormal", "Normal"))
    cat("\n")
    tb <- table(truth = temp$htype, pred = temp$ctype)
    print(tb)
    
    tb2 <- table(truth = temp$htype, pred = temp$ctype2)
    print(tb2)
    tb3 <- table(truth = temp$htype, pred = temp$ctype3)
    print(tb3)
  }

}

```


### 4.3 Precision

Find cutoff of functional scores with precision of 0.95, 0.98 and 0.99 for both StopGain and Synonymous using our data and Clinvar subsets and HDR subsets if there are any. 
```{r prec_fun}

myprec <- function(dt, typecolumn, casen, controln, precisions = c(0.9, 0.95, 0.98, 0.99)){
  temp <- dt
  temp$type <- pull(dt, typecolumn)
  score = c(temp$functional.score[which(temp$type == casen)], temp$functional.score[which(temp$type == controln)])
  label = c(temp$type[which(temp$type == casen)], temp$type[temp$type == controln])
  label = factor(label, levels = c(casen, controln))
  weights = c(1, 1)
  
  cutoffs = precisionBasedThreshold(score, label, weights, precisions)
  
  p <- temp %>%
    ggplot( aes(x=functional.score, fill=type)) +
    geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity', binwidth = 0.1) +
    geom_vline(xintercept=c(cutoffs[,casen]), linetype="dashed", color = "red") + ## upper stopgain
    geom_vline(xintercept=c(cutoffs[, controln]), linetype="dashed", color = "green") + ## lower synonymous
    scale_fill_manual(values=c("#69b3a2", "red")) +  ggtitle ( "functional.score with precision cutoffs(weight 1)")
  ggplotly(p)
  
  weights = c(1, sum(temp$type ==casen) / sum(temp$type == controln))
  cutoffs2 = precisionBasedThreshold(score, label, weights, precisions)
  
  p <- temp %>%
    ggplot( aes(x=functional.score, fill=type)) +
    geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity', binwidth = 0.1) +
    geom_vline(xintercept=c(cutoffs2[,casen]), linetype="dashed", color = "red") + ## upper stopgain
    geom_vline(xintercept=c(cutoffs2[, controln]), linetype="dashed", color = "green") + ## lower synonymous
    scale_fill_manual(values=c("#69b3a2", "red")) +  ggtitle ( "functional.score with precision cutoffs (data weight)")
  ggplotly(p)
  
  weights = c(1, 3*sum(temp$type ==casen) / sum(temp$type == controln))
  cutoffs3 = precisionBasedThreshold(score, label, weights, precisions)
  
  p <- temp %>%
    ggplot( aes(x=functional.score, fill=type)) +
    geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity', binwidth = 0.1) +
    geom_vline(xintercept=c(cutoffs3[,casen]), linetype="dashed", color = "red") + ## upper stopgain
    geom_vline(xintercept=c(cutoffs3[, controln]), linetype="dashed", color = "green") + ## lower synonymous
    scale_fill_manual(values=c("#69b3a2", "red")) +  ggtitle ( "functional.score with precision cutoffs (alpha weight)")
  ggplotly(p)
  
  return(list(cutoffs = cutoffs, cutoffs2 = cutoffs2, cutoffs3 = cutoffs3))
}

```

#### Data driven

Based on distribution from StopGain and synonymous non splicing region (spliceR == FALSE), excluding variants on chunling's list, we also build precision cutoffs.

```{r prec_data, include=TRUE, fig.width=9, fig.height=5, message=FALSE, warning=FALSE}

temp <- D14vsLib_fs.2 %>% filter(!is.na(functional.score)) %>% filter(EventType %in% c("StopGain", "Synonymous")) %>% 
  arrange(functional.score) %>% filter(spliceR== FALSE) 
if(!is.null(params$vexclude)){
    temp <- temp %>%  filter(!uPOS %in% Varaints_to_be_excluded_when_use_MAVE_internal_control_to_build_model$uPOS)  ## remove variants on chunling's list
}

cfs <- myprec(temp, typecolumn = "EventType", casen = "StopGain", controln = "Synonymous")
```


for precision cutoffs with weight ratio of 1, 1.
```{r}
print(cfs$cutoffs)
```
for precision cutoffs with weight ratio of data (1, `r sum(temp$EventType == "StopGain") / sum(temp$EventType == "Synonymous")`).
```{r}
print(cfs$cutoffs2)
```
for precision cutoffs with weight ratio of data (1, `r 3*sum(temp$EventType == "StopGain") / sum(temp$EventType == "Synonymous")`).
```{r}
print(cfs$cutoffs3)
```

Applying the cutoffs on three types together (MAVE data).

- cutoffs with weight ratio of 1. 
Note: Splice regions are included in Missense, not stopgain nor synonymous



```{r prec_data2, include=TRUE, fig.width=9, fig.height=5, message=FALSE, warning=FALSE}
temp <- D14vsLib_fs.2 %>% filter(EventType %in% c("Synonymous", "StopGain")) %>% filter(spliceR == FALSE) 

temp2 <- D14vsLib_fs.2 %>% filter(EventType %in% c("Missense")) 

temp <- rbind(temp, temp2)
precisions = c(0.9, 0.95, 0.98, 0.99)
for(ri in precisions){
  cat("\n", "precision ", ri, "\n")
  temp$ctype.p <- ifelse(temp$functional.score>cfs$cutoffs[as.character(ri), "Synonymous"], "neutural", 
                     ifelse(temp$functional.score>cfs$cutoffs[as.character(ri), "StopGain"], "intermediate", "pathogenic"))
  temp$ctype.p <- factor(temp$ctype.p, levels = c("pathogenic", "intermediate", "neutural"))
  temp$EventType <-factor(temp$EventType, levels = c("StopGain", "Missense", "Synonymous"))

  tb <- table(truth = temp$EventType, pred = temp$ctype.p)
  print(tb)
}
```

- cutoffs with weight ratio of data ratio. 
Note: Splice regions are included in Missense, not stopgain nor synonymous

```{r prec_data2.1, include=TRUE, fig.width=9, fig.height=5, message=FALSE, warning=FALSE}
for(ri in precisions){
  cat("\n", "precision ", ri, "\n")
  temp$ctype.p <- ifelse(temp$functional.score>cfs$cutoffs2[as.character(ri), "Synonymous"], "neutural", 
                     ifelse(temp$functional.score>cfs$cutoffs2[as.character(ri), "StopGain"], "intermediate", "pathogenic"))
  temp$ctype.p <- factor(temp$ctype.p, levels = c("pathogenic", "intermediate", "neutural"))
  temp$EventType <-factor(temp$EventType, levels = c("StopGain", "Missense", "Synonymous"))

  tb <- table(truth = temp$EventType, pred = temp$ctype.p)
  print(tb)
}
```

- cutoffs with weight ratio of alpha ratio. 
Note: Splice regions are included in Missense, not stopgain nor synonymous

```{r prec_data2.1a, include=TRUE, fig.width=9, fig.height=5, message=FALSE, warning=FALSE}
for(ri in precisions){
  cat("\n", "precision ", ri, "\n")
  temp$ctype.p <- ifelse(temp$functional.score>cfs$cutoffs3[as.character(ri), "Synonymous"], "neutural", 
                     ifelse(temp$functional.score>cfs$cutoffs3[as.character(ri), "StopGain"], "intermediate", "pathogenic"))
  temp$ctype.p <- factor(temp$ctype.p, levels = c("pathogenic", "intermediate", "neutural"))
  temp$EventType <-factor(temp$EventType, levels = c("StopGain", "Missense", "Synonymous"))

  tb <- table(truth = temp$EventType, pred = temp$ctype.p)
  print(tb)
}
```

```{r}

if(!is.null(params$clinvarfile)){
  
  cat("\n")
    
  cat("Applying the cutoffs on Clinvar data excluding spliceR.\n")
  
  cat("- cutoffs with weight ratio of 1.\n")

}

```



```{r prec_data2c, include=TRUE, fig.width=9, fig.height=5, message=FALSE, warning=FALSE}

if(!is.null(params$clinvarfile)){
  
temp <- D14vsLib_fs.2 %>% filter(type %in% c("Benign", "Pathogenic")) %>% filter(spliceR == FALSE) 

for(ri in precisions){
  cat("\n", "precision ", ri, "\n")
  temp$ctype.p <- ifelse(temp$functional.score>cfs$cutoffs[as.character(ri), "Synonymous"], "neutural", 
                     ifelse(temp$functional.score>cfs$cutoffs[as.character(ri), "StopGain"], "intermediate", "pathogenic"))
  temp$ctype.p <- factor(temp$ctype.p, levels = c("pathogenic", "intermediate", "neutural"))
  temp$EventType <-factor(temp$EventType, levels = c("StopGain", "Missense", "Synonymous"))

  tb <- table(truth = temp$EventType, pred = temp$ctype.p)
  print(tb)
}
}

```

`r cat(ifelse(!is.null(params$clinvarfile), "- cutoffs with weight ratio of data ratio. \n", "\n"))`

```{r prec_data2.1cc, include=TRUE, fig.width=9, fig.height=5, message=FALSE, warning=FALSE}

if(!is.null(params$clinvarfile)){
for(ri in precisions){
  cat("\n", "precision ", ri, "\n")
  temp$ctype.p <- ifelse(temp$functional.score>cfs$cutoffs2[as.character(ri), "Synonymous"], "neutural", 
                     ifelse(temp$functional.score>cfs$cutoffs2[as.character(ri), "StopGain"], "intermediate", "pathogenic"))
  temp$ctype.p <- factor(temp$ctype.p, levels = c("pathogenic", "intermediate", "neutural"))
  temp$EventType <-factor(temp$EventType, levels = c("StopGain", "Missense", "Synonymous"))

  tb <- table(truth = temp$EventType, pred = temp$ctype.p)
  print(tb)
}
}

```

`r cat(ifelse(!is.null(params$clinvarfile), "- cutoffs with weight ratio of alpha ratio.  \n", "\n"))`



```{r prec_data2.1acc, include=TRUE, fig.width=9, fig.height=5, message=FALSE, warning=FALSE}
if(!is.null(params$clinvarfile)){
for(ri in precisions){
  cat("\n", "precision ", ri, "\n")
  temp$ctype.p <- ifelse(temp$functional.score>cfs$cutoffs3[as.character(ri), "Synonymous"], "neutural", 
                     ifelse(temp$functional.score>cfs$cutoffs3[as.character(ri), "StopGain"], "intermediate", "pathogenic"))
  temp$ctype.p <- factor(temp$ctype.p, levels = c("pathogenic", "intermediate", "neutural"))
  temp$EventType <-factor(temp$EventType, levels = c("StopGain", "Missense", "Synonymous"))

  tb <- table(truth = temp$EventType, pred = temp$ctype.p)
  print(tb)
}
}

```

```{r}

if(!is.null(params$clinvarfile)){
  cat("\n")
  
  cat("#### ClinVar Data driven\n")
  
  cat("Based on distribution from Clinvar (pathogentic) and (benign), we also build precision model. \n")
  cat("Exlcude spliceR and variants on chunling's list.\n")
}
```


```{r prec.clinvar, include=TRUE, fig.width=9, fig.height=5, message=FALSE, warning=FALSE}

if(!is.null(params$clinvarfile)){
temp <- D14vsLib_fs.2 %>% filter(!is.na(functional.score)) %>% arrange(functional.score) %>% filter(type %in% c("Pathogenic", "Benign")) %>% filter(spliceR == FALSE) 
if(!is.null(params$vexclude)){
    temp <- temp %>%  filter(!uPOS %in% Varaints_to_be_excluded_when_use_MAVE_internal_control_to_build_model$uPOS)  ## remove variants on chunling's list
}

cfs.c <- myprec(temp, typecolumn = "type", casen = "Pathogenic", controln = "Benign")

cat("  for precision cutoffs with weight ratio of 1, 1.\n")
print(cfs.c$cutoffs)
cat("\n")
cat(paste0("for precision cutoffs with weight ratio of data (1,", sum(temp$type == "Pathogenic") / sum(temp$type == "Benign")))
cat("\n")
print(cfs.c$cutoffs2)

cat(paste0("for precision cutoffs with weight ratio of data (1, ", 3*sum(temp$type == "Pathogenic") / sum(temp$type == "Benign")))
cat("\n")
print(cfs.c$cutoffs3)
}

```

```{r}

if(!is.null(params$clinvarfile)){
  cat("\n")
  cat("Applying the cutoffs on three types together for MAVE data.\n")
  
  cat("- cutoffs with weight ratio of 1. \n")
  cat("Note: Splice regions are included in Missense, not stopgain nor synonymous\n")

}
```


```{r prec_data2.c, include=TRUE, fig.width=9, fig.height=5, message=FALSE, warning=FALSE}
if(!is.null(params$clinvarfile)){
temp <- D14vsLib_fs.2 %>% filter(EventType %in% c("Synonymous", "StopGain")) %>% filter(spliceR == FALSE) 

temp2 <- D14vsLib_fs.2 %>% filter(EventType %in% c("Missense")) 

temp <- rbind(temp, temp2)

for(ri in precisions){
  cat("\n", "precision ", ri, "\n")
  temp$ctype.p <- ifelse(temp$functional.score>cfs.c$cutoffs[as.character(ri), "Benign"], "neutural", 
                     ifelse(temp$functional.score>cfs.c$cutoffs[as.character(ri), "Pathogenic"], "intermediate", "pathogenic"))
  temp$ctype.p <- factor(temp$ctype.p, levels = c("pathogenic", "intermediate", "neutural"))
  temp$EventType <-factor(temp$EventType, levels = c("StopGain", "Missense", "Synonymous"))

  tb <- table(truth = temp$EventType, pred = temp$ctype.p)
  print(tb)
}
}
```

```{r}
if(!is.null(params$clinvarfile)){
   cat("\n")
  cat("- cutoffs with weight ratio of data ratio. \n")
    
  cat("Note: Splice regions are included in Missense, not stopgain nor synonymous\n")

}
```


```{r prec_data2.1c, include=TRUE, fig.width=9, fig.height=5, message=FALSE, warning=FALSE}
if(!is.null(params$clinvarfile)){
for(ri in precisions){
  cat("\n", "precision ", ri, "\n")
  temp$ctype.p <- ifelse(temp$functional.score>cfs.c$cutoffs2[as.character(ri), "Benign"], "neutural", 
                     ifelse(temp$functional.score>cfs.c$cutoffs2[as.character(ri), "Pathogenic"], "intermediate", "pathogenic"))
  temp$ctype.p <- factor(temp$ctype.p, levels = c("pathogenic", "intermediate", "neutural"))
  temp$EventType <-factor(temp$EventType, levels = c("StopGain", "Missense", "Synonymous"))

  tb <- table(truth = temp$EventType, pred = temp$ctype.p)
  print(tb)
}
}

```

```{r}
if(!is.null(params$clinvarfile)){
   cat("\n")
cat("- cutoffs with weight ratio of alpha ratio. \n")
cat("Note: Splice regions are included in Missense, not stopgain nor synonymous \n")

}
```


```{r prec_data2.1ca, include=TRUE, fig.width=9, fig.height=5, message=FALSE, warning=FALSE}
if(!is.null(params$clinvarfile)){
for(ri in precisions){
  cat("\n", "precision ", ri, "\n")
  temp$ctype.p <- ifelse(temp$functional.score>cfs.c$cutoffs3[as.character(ri), "Benign"], "neutural", 
                     ifelse(temp$functional.score>cfs.c$cutoffs3[as.character(ri), "Pathogenic"], "intermediate", "pathogenic"))
  temp$ctype.p <- factor(temp$ctype.p, levels = c("pathogenic", "intermediate", "neutural"))
  temp$EventType <-factor(temp$EventType, levels = c("StopGain", "Missense", "Synonymous"))

  tb <- table(truth = temp$EventType, pred = temp$ctype.p)
  print(tb)
}
}
```

```{r}
if(!is.null(params$clinvarfile)){
 cat("\n")
  cat("Applying the cutoffs on Clinvar data\n")
  
  cat("- cutoffs with weight ratio of 1. \n")
}

```

    

```{r prec_data2.ccc, include=TRUE, fig.width=9, fig.height=5, message=FALSE, warning=FALSE}
if(!is.null(params$clinvarfile)){
temp <- D14vsLib_fs.2 %>% filter(type %in% c("Benign", "Pathogenic")) %>% filter(spliceR == FALSE) 

for(ri in precisions){
  cat("\n", "precision ", ri, "\n")
  temp$ctype.p <- ifelse(temp$functional.score>cfs.c$cutoffs[as.character(ri), "Benign"], "neutural", 
                     ifelse(temp$functional.score>cfs.c$cutoffs[as.character(ri), "Pathogenic"], "intermediate", "pathogenic"))
  temp$ctype.p <- factor(temp$ctype.p, levels = c("pathogenic", "intermediate", "neutural"))
  temp$type <-factor(temp$type, levels = c("Pathogenic", "Benign"))

  tb <- table(truth = temp$type, pred = temp$ctype.p)
  print(tb)
}
}
```

`r cat(ifelse(!is.null(params$clinvarfile), "- cutoffs with weight ratio of data ratio. \n", "\n"))`

```{r prec_data2.1ccc, include=TRUE, fig.width=9, fig.height=5, message=FALSE, warning=FALSE}
if(!is.null(params$clinvarfile)){
for(ri in precisions){
  cat("\n", "precision ", ri, "\n")
  temp$ctype.p <- ifelse(temp$functional.score>cfs.c$cutoffs2[as.character(ri), "Benign"], "neutural", 
                     ifelse(temp$functional.score>cfs.c$cutoffs2[as.character(ri), "Pathogenic"], "intermediate", "pathogenic"))
  temp$ctype.p <- factor(temp$ctype.p, levels = c("pathogenic", "intermediate", "neutural"))
  temp$type <-factor(temp$type, levels = c("Pathogenic", "Benign"))

  tb <- table(truth = temp$type, pred = temp$ctype.p)
  print(tb)
}
}

```

`r cat(ifelse(!is.null(params$clinvarfile), "- cutoffs with weight ratio of alpha ratio. \n", "\n"))`

```{r prec_data2.1cacc, include=TRUE, fig.width=9, fig.height=5, message=FALSE, warning=FALSE}
if(!is.null(params$clinvarfile)){
for(ri in precisions){
  cat("\n", "precision ", ri, "\n")
  temp$ctype.p <- ifelse(temp$functional.score>cfs.c$cutoffs3[as.character(ri), "Benign"], "neutural", 
                     ifelse(temp$functional.score>cfs.c$cutoffs3[as.character(ri), "Pathogenic"], "intermediate", "pathogenic"))
  temp$ctype.p <- factor(temp$ctype.p, levels = c("pathogenic", "intermediate", "neutural"))
  temp$type <-factor(temp$type, levels = c("Pathogenic", "Benign"))

  tb <- table(truth = temp$type, pred = temp$ctype.p)
  print(tb)
}
}
```




```{r, results = 'asis',  include=TRUE, fig.width=9, fig.height=5, message=FALSE, warning=FALSE}

if(!is.null(params$HDRfile)){

  cat("\n")
  cat("#### HDR Data driven\n")

  cat("\n")
  cat("Based on distribution from HDR data (Abnormal) and (Normal), we also build precision model. Exlcude spliceR and variants on chunling's list.\n")
  temp <- D14vsLib_fs.2 %>% filter(!is.na(functional.score)) %>% arrange(functional.score) %>% filter(htype %in% c("Abnormal", "Normal")) %>% filter(spliceR == FALSE) 
  if(!is.null(params$vexclude)){
    temp <- temp %>%  filter(!uPOS %in% Varaints_to_be_excluded_when_use_MAVE_internal_control_to_build_model$uPOS)  ## remove variants on chunling's list
  }
  
  cfs.h <- myprec(temp, typecolumn = "htype", casen = "Abnormal", controln = "Normal")
}
```


`r print(ifelse(!is.null(params$HDRfile), " for precision cutoffs with weight ratio of 1, 1. ", ""))`

```{r}

if(!is.null(params$HDRfile)){
  print(cfs.h$cutoffs)
}
```


`r ifelse(!is.null(params$HDRfile), paste0(" for precision cutoffs with weight ratio of data: 1, ", sum(temp$htype == "Abnormal") / sum(temp$htype == "Normal"), " .."), "")` 

```{r}

if(!is.null(params$HDRfile)){
  print(cfs.h$cutoffs2)
}
```

`r ifelse(!is.null(params$HDRfile),  paste0(" for precision cutoffs with weight ratio of data: 1, ", 3*sum(temp$htype == "Abnormal") / sum(temp$htype == "Normal"), " .."), "")`

```{r}

if(!is.null(params$HDRfile)){
  print(cfs.h$cutoffs3)
}
```

`r ifelse(!is.null(params$HDRfile), " Applying the cutoffs on three types together for MAVE data. ", "")`

`r ifelse(!is.null(params$HDRfile), " - cutoffs with weight ratio of 1. ", "")`

`r ifelse(!is.null(params$HDRfile), " Note: Splice regions are included in Missense, not stopgain nor synonymous. ", "")`


```{r}

if(!is.null(params$HDRfile)){
  temp <- D14vsLib_fs.2 %>% filter(EventType %in% c("Synonymous", "StopGain")) %>% filter(spliceR == FALSE) 

  temp2 <- D14vsLib_fs.2 %>% filter(EventType %in% c("Missense")) 

  temp <- rbind(temp, temp2)

  for(ri in precisions){
    cat("\n", "precision ", ri, "\n")
    temp$ctype.p <- ifelse(temp$functional.score>cfs.h$cutoffs[as.character(ri), "Normal"], "neutural", 
                       ifelse(temp$functional.score>cfs.h$cutoffs[as.character(ri), "Abnormal"], "intermediate", "pathogenic"))
    temp$ctype.p <- factor(temp$ctype.p, levels = c("pathogenic", "intermediate", "neutural"))
    temp$EventType <-factor(temp$EventType, levels = c("StopGain", "Missense", "Synonymous"))
  
    tb <- table(truth = temp$EventType, pred = temp$ctype.p)
    print(tb)
  }
}

```

`r ifelse(!is.null(params$HDRfile), " - cutoffs with weight ratio of data ratio.. ", "")` 
`r ifelse(!is.null(params$HDRfile), " Note: Splice regions are included in Missense, not stopgain nor synonymous. ", "")` 

```{r}

if(!is.null(params$HDRfile)){
  for(ri in precisions){
    cat("\n", "precision ", ri, "\n")
    temp$ctype.p <- ifelse(temp$functional.score>cfs.h$cutoffs2[as.character(ri), "Normal"], "neutural", 
                       ifelse(temp$functional.score>cfs.h$cutoffs2[as.character(ri), "Abnormal"], "intermediate", "pathogenic"))
    temp$ctype.p <- factor(temp$ctype.p, levels = c("pathogenic", "intermediate", "neutural"))
    temp$EventType <-factor(temp$EventType, levels = c("StopGain", "Missense", "Synonymous"))
  
    tb <- table(truth = temp$EventType, pred = temp$ctype.p)
    print(tb)
  }
}
```

`r ifelse(!is.null(params$HDRfile), " - cutoffs with weight ratio of alpha ratio.. ", "")` 
`r ifelse(!is.null(params$HDRfile), " Note: Splice regions are included in Missense, not stopgain nor synonymous. ", "")` 

  
```{r}

if(!is.null(params$HDRfile)){
  for(ri in precisions){
    cat("\n", "precision ", ri, "\n")
    temp$ctype.p <- ifelse(temp$functional.score>cfs.h$cutoffs3[as.character(ri), "Normal"], "neutural", 
                       ifelse(temp$functional.score>cfs.h$cutoffs3[as.character(ri), "Abnormal"], "intermediate", "pathogenic"))
    temp$ctype.p <- factor(temp$ctype.p, levels = c("pathogenic", "intermediate", "neutural"))
    temp$EventType <-factor(temp$EventType, levels = c("StopGain", "Missense", "Synonymous"))
  
    tb <- table(truth = temp$EventType, pred = temp$ctype.p)
    print(tb)
  }
}
```


`r ifelse(!is.null(params$HDRfile), "Applying the cutoffs on HDR data. ", "")` 
`r ifelse(!is.null(params$HDRfile), " - cutoffs with weight ratio of 1. ", "")` 
`r ifelse(!is.null(params$HDRfile), " Note: Splice regions are included in Missense, not stopgain nor synonymous. ", "")` 

```{r}

if(!is.null(params$HDRfile)){
  temp <- D14vsLib_fs.2 %>% filter(htype %in% c("Abnormal", "Normal")) %>% filter(spliceR == FALSE) 

  for(ri in precisions){
    cat("\n", "precision ", ri, "\n")
    temp$ctype.p <- ifelse(temp$functional.score>cfs.h$cutoffs[as.character(ri), "Normal"], "neutural", 
                       ifelse(temp$functional.score>cfs.h$cutoffs[as.character(ri), "Abnormal"], "intermediate", "pathogenic"))
    temp$ctype.p <- factor(temp$ctype.p, levels = c("pathogenic", "intermediate", "neutural"))
  temp$type <-factor(temp$htype, levels = c("Abnormal", "Normal"))

  tb <- table(truth = temp$type, pred = temp$ctype.p)
    print(tb)
  }
  
}
```

`r ifelse(!is.null(params$HDRfile), " - cutoffs with weight ratio of data ratio. ", "")` 
`r ifelse(!is.null(params$HDRfile), " Note: Splice regions are included in Missense, not stopgain nor synonymous. ", "")` 
 
```{r}

if(!is.null(params$HDRfile)){ 
  for(ri in precisions){
    cat("\n", "precision ", ri, "\n")
    temp$ctype.p <- ifelse(temp$functional.score>cfs.h$cutoffs2[as.character(ri), "Normal"], "neutural", 
                       ifelse(temp$functional.score>cfs.h$cutoffs2[as.character(ri), "Abnormal"], "intermediate", "pathogenic"))
    temp$ctype.p <- factor(temp$ctype.p, levels = c("pathogenic", "intermediate", "neutural"))
      temp$type <-factor(temp$htype, levels = c("Abnormal", "Normal"))

  tb <- table(truth = temp$type, pred = temp$ctype.p)
    print(tb)
    
  }
  
}
```

`r ifelse(!is.null(params$HDRfile), " - cutoffs with weight alpha ratio. ", "")` 
`r ifelse(!is.null(params$HDRfile), " Note: Splice regions are included in Missense, not stopgain nor synonymous. ", "")` 

```{r}

if(!is.null(params$HDRfile)){ 
  for(ri in precisions){
    cat("\n", "precision ", ri, "\n")
    temp$ctype.p <- ifelse(temp$functional.score>cfs.h$cutoffs3[as.character(ri), "Normal"], "neutural", 
                       ifelse(temp$functional.score>cfs.h$cutoffs3[as.character(ri), "Abnormal"], "intermediate", "pathogenic"))
    temp$ctype.p <- factor(temp$ctype.p, levels = c("pathogenic", "intermediate", "neutural"))
      temp$type <-factor(temp$htype, levels = c("Abnormal", "Normal"))

  tb <- table(truth = temp$type, pred = temp$ctype.p)
    print(tb)
  }
}

```




```{r}
## output merge with raw D14_v_lib
out <- merge(D14vsLib_fs.2.av, dt2.av.temp) 
if(nmax == 6){
  out <- out %>% rowwise() %>%  dplyr::select(-E1, -E2, -E3, -E4, -E5, -E6) %>% relocate(functional.score, .after = R6_ce_norm)
}else if(nmax == 5){
  out <- out %>% rowwise() %>%  dplyr::select(-E1, -E2, -E3, -E4, -E5) %>% relocate(functional.score, .after = R5_ce_norm)
}else if(nmax == 4){
  out <- out %>% rowwise() %>%  dplyr::select(-E1, -E2, -E3, -E4) %>% relocate(functional.score, .after = R4_ce_norm)
}else if(nmax == 3){
  out <- out %>% rowwise() %>%  dplyr::select(-E1, -E2, -E3) %>% relocate(functional.score, .after = R3_ce_norm)
}else if(nmax == 2){
  out <- out %>% rowwise() %>%  dplyr::select(-E1, -E2) %>% relocate(functional.score, .after = R2_ce_norm)
} else {
  out <- out %>% rowwise() %>%  dplyr::select(-E1) %>% relocate(functional.score, .after = R1_ce_norm)
}

## output merge with raw counts
## add the mm cutoffs
## data
Sout <- data.frame(m95.sg.pd = mm0$f95, m95.sn.pd = mm0$f5,
                   m98.sg.pd = mm0$f98, m98.sn.pd = mm0$f2,
                   m99.sg.pd = mm0$f99, m99.sn.pd = mm0$f1,
                   m95.sg.p01= mm0.p01$f95, m95.sn.p01 = mm0.p01$f5,
                   m98.sg.p01= mm0.p01$f98, m98.sn.p01 = mm0.p01$f2,
                   m99.sg.p01= mm0.p01$f99, m99.sn.p01 = mm0.p01$f1,
                   m95.sg.p02= mm0.p02$f95, m95.sn.p02 = mm0.p02$f5,
                   m98.sg.p02= mm0.p02$f98, m98.sn.p02 = mm0.p02$f2,
                   m99.sg.p02= mm0.p02$f99, m99.sn.p02 = mm0.p02$f1)
if(!is.null(params$clinvarfile)){         
                   ## clinvar
Sout <- Sout %>% mutate(
                   
                   m95.sg.Cpd = mm0.c$f95, m95.sn.Cpd = mm0.c$f5,
                   m98.sg.Cpd = mm0.c$f98, m98.sn.Cpd = mm0.c$f2,
                   m99.sg.Cpd = mm0.c$f99, m99.sn.Cpd = mm0.c$f1,
                   m95.sg.Cp01= mm0.c.p01$f95, m95.sn.Cp01 = mm0.c.p01$f5,
                   m98.sg.Cp01= mm0.c.p01$f98, m98.sn.Cp01 = mm0.c.p01$f2,
                   m99.sg.Cp01= mm0.c.p01$f99, m99.sn.Cp01 = mm0.c.p01$f1,
                   m95.sg.Cp02= mm0.c.p02$f95, m95.sn.Cp02 = mm0.c.p02$f5,
                   m98.sg.Cp02= mm0.c.p02$f98, m98.sn.Cp02 = mm0.c.p02$f2,
                   m99.sg.Cp02= mm0.c.p02$f99, m99.sn.Cp02 = mm0.c.p02$f1)
} else{
  Sout <- Sout %>% mutate(m95.sg.Cpd = NA, m95.sn.Cpd = NA,
                   m98.sg.Cpd = NA, m98.sn.Cpd = NA,
                   m99.sg.Cpd = NA, m99.sn.Cpd = NA,
                   m95.sg.Cp01= NA, m95.sn.Cp01 = NA,
                   m98.sg.Cp01= NA, m98.sn.Cp01 = NA,
                   m99.sg.Cp01= NA, m99.sn.Cp01 = NA,
                   m95.sg.Cp02= NA, m95.sn.Cp02 = NA,
                   m98.sg.Cp02= NA, m98.sn.Cp02 = NA,
                   m99.sg.Cp02= NA, m99.sn.Cp02 = NA)
}
if(!is.null(params$HDRfile)){ 
  Sout <- Sout %>% mutate(
                   m95.sg.Hpd = mm0.h$f95, m95.sn.Hpd = mm0.h$f5,
                   m98.sg.Hpd = mm0.h$f98, m98.sn.Hpd = mm0.h$f2,
                   m99.sg.Hpd = mm0.h$f99, m99.sn.Hpd = mm0.h$f1,
                   m95.sg.Hp01= mm0.h.p01$f95, m95.sn.Hp01 = mm0.h.p01$f5,
                   m98.sg.Hp01= mm0.h.p01$f98, m98.sn.Hp01 = mm0.h.p01$f2,
                   m99.sg.Hp01= mm0.h.p01$f99, m99.sn.Hp01 = mm0.h.p01$f1,
                   m95.sg.Hp02= mm0.h.p02$f95, m95.sn.Hp02 = mm0.h.p02$f5,
                   m98.sg.Hp02= mm0.h.p02$f98, m98.sn.Hp02 = mm0.h.p02$f2,
                   m99.sg.Hp02= mm0.h.p02$f99, m99.sn.Hp02 = mm0.h.p02$f1)
}

                   ## add the precision cutoffs
 Sout <- Sout %>% mutate(
                   p90.sg.w1 = cfs$cutoffs["0.9", "StopGain"], p90.sn.w1 = cfs$cutoffs["0.9", "Synonymous"], 
                   p95.sg.w1 = cfs$cutoffs["0.95", "StopGain"], p95.sn.w1 = cfs$cutoffs["0.95", "Synonymous"], 
                   p98.sg.w1 = cfs$cutoffs["0.98", "StopGain"], p98.sn.w1 = cfs$cutoffs["0.98", "Synonymous"], 
                   p99.sg.w1 = cfs$cutoffs["0.99", "StopGain"], p99.sn.w1 = cfs$cutoffs["0.99", "Synonymous"], 
                   
                   p90.sg.wd = cfs$cutoffs2["0.9", "StopGain"], p90.sn.wd = cfs$cutoffs2["0.9", "Synonymous"], 
                   p95.sg.wd = cfs$cutoffs2["0.95", "StopGain"], p95.sn.wd = cfs$cutoffs2["0.95", "Synonymous"], 
                   p98.sg.wd = cfs$cutoffs2["0.98", "StopGain"], p98.sn.wd = cfs$cutoffs2["0.98", "Synonymous"], 
                   p99.sg.wd = cfs$cutoffs2["0.99", "StopGain"], p99.sn.wd = cfs$cutoffs2["0.99", "Synonymous"], 
                   
                   p90.sg.wa = cfs$cutoffs3["0.9", "StopGain"], p90.sn.wa = cfs$cutoffs3["0.9", "Synonymous"], 
                   p95.sg.wa = cfs$cutoffs3["0.95", "StopGain"], p95.sn.wa = cfs$cutoffs3["0.95", "Synonymous"], 
                   p98.sg.wa = cfs$cutoffs3["0.98", "StopGain"], p98.sn.wa = cfs$cutoffs3["0.98", "Synonymous"], 
                   p99.sg.wa = cfs$cutoffs3["0.99", "StopGain"], p99.sn.wa = cfs$cutoffs3["0.99", "Synonymous"])
 
if(!is.null(params$clinvarfile)){         
                   ## clinvar
Sout <- Sout %>% mutate(
  
                  p90.sg.w1.c = cfs.c$cutoffs["0.9", "Pathogenic"], p90.sn.w1.c = cfs.c$cutoffs["0.9", "Benign"], 
                   p95.sg.w1.c = cfs.c$cutoffs["0.95", "Pathogenic"], p95.sn.w1.c = cfs.c$cutoffs["0.95", "Benign"], 
                   p98.sg.w1.c = cfs.c$cutoffs["0.98", "Pathogenic"], p98.sn.w1.c = cfs.c$cutoffs["0.98", "Benign"], 
                   p99.sg.w1.c = cfs.c$cutoffs["0.99", "Pathogenic"], p99.sn.w1.c = cfs.c$cutoffs["0.99", "Benign"], 
                   
                   p90.sg.wd.c = cfs.c$cutoffs2["0.9", "Pathogenic"], p90.sn.wd.c = cfs.c$cutoffs2["0.9", "Benign"], 
                   p95.sg.wd.c = cfs.c$cutoffs2["0.95", "Pathogenic"], p95.sn.wd.c = cfs.c$cutoffs2["0.95", "Benign"], 
                   p98.sg.wd.c = cfs.c$cutoffs2["0.98","Pathogenic"], p98.sn.wd.c = cfs.c$cutoffs2["0.98", "Benign"], 
                   p99.sg.wd.c = cfs.c$cutoffs2["0.99", "Pathogenic"], p99.sn.wd.c = cfs.c$cutoffs2["0.99", "Benign"], 
                   
                   p90.sg.wa.c = cfs.c$cutoffs3["0.9", "Pathogenic"], p90.sn.wa.c = cfs.c$cutoffs3["0.9", "Benign"], 
                   p95.sg.wa.c = cfs.c$cutoffs3["0.95", "Pathogenic"], p95.sn.wa.c = cfs.c$cutoffs3["0.95", "Benign"], 
                   p98.sg.wa.c = cfs.c$cutoffs3["0.98","Pathogenic"], p98.sn.wa.c = cfs.c$cutoffs3["0.98", "Benign"], 
                   p99.sg.wa.c = cfs.c$cutoffs3["0.99", "Pathogenic"], p99.sn.wa.c = cfs.c$cutoffs3["0.99", "Benign"]
  
)
} else{
  Sout <- Sout %>% mutate(
                   p90.sg.w1.c = NA, p90.sn.w1.c = NA, 
                   p95.sg.w1.c = NA, p95.sn.w1.c = NA, 
                   p98.sg.w1.c = NA, p98.sn.w1.c = NA, 
                   p99.sg.w1.c = NA, p99.sn.w1.c = NA, 
                   
                   p90.sg.wd.c = NA, p90.sn.wd.c = NA, 
                   p95.sg.wd.c = NA, p95.sn.wd.c = NA, 
                   p98.sg.wd.c = NA, p98.sn.wd.c = NA, 
                   p99.sg.wd.c = NA, p99.sn.wd.c = NA, 
                   
                   p90.sg.wa.c = NA, p90.sn.wa.c = NA, 
                   p95.sg.wa.c = NA, p95.sn.wa.c = NA, 
                   p98.sg.wa.c = NA, p98.sn.wa.c = NA, 
                   p99.sg.wa.c = NA, p99.sn.wa.c = NA)
}


## HDR
if(!is.null(params$HDRfile)){
  Sout$p90.sg.w1.h <- cfs.h$cutoffs["0.9", "Abnormal"]; Sout$p90.sn.w1.h <- cfs.h$cutoffs["0.9", "Normal"]; 
  Sout$p95.sg.w1.h <- cfs.h$cutoffs["0.95", "Abnormal"]; Sout$p95.sn.w1.h <- cfs.h$cutoffs["0.95", "Normal"]; 
  Sout$p98.sg.w1.h <- cfs.h$cutoffs["0.98", "Abnormal"]; Sout$p98.sn.w1.h <- cfs.h$cutoffs["0.98", "Normal"]; 
  Sout$p99.sg.w1.h <- cfs.h$cutoffs["0.99", "Abnormal"]; Sout$p99.sn.w1.h <- cfs.h$cutoffs["0.99", "Normal"]; 
  
  Sout$p90.sg.wd.h <- cfs.h$cutoffs2["0.9", "Abnormal"]; Sout$p90.sn.wd.h <- cfs.h$cutoffs2["0.9", "Normal"]; 
  Sout$p95.sg.wd.h <- cfs.h$cutoffs2["0.95", "Abnormal"]; Sout$p95.sn.wd.h <- cfs.h$cutoffs2["0.95", "Normal"]; 
  Sout$p98.sg.wd.h <- cfs.h$cutoffs2["0.98","Abnormal"]; Sout$p98.sn.wd.h <- cfs.h$cutoffs2["0.98", "Normal"]; 
  Sout$p99.sg.wd.h <- cfs.h$cutoffs2["0.99", "Abnormal"]; Sout$p99.sn.wd.h <- cfs.h$cutoffs2["0.99", "Normal"]; 
  
  Sout$p90.sg.wa.h <- cfs.h$cutoffs3["0.9", "Abnormal"]; Sout$p90.sn.wa.h <- cfs.h$cutoffs3["0.9", "Normal"]; 
  Sout$p95.sg.wa.h <- cfs.h$cutoffs3["0.95", "Abnormal"]; Sout$p95.sn.wa.h <- cfs.h$cutoffs3["0.95", "Normal"]; 
  Sout$p98.sg.wa.h <- cfs.h$cutoffs3["0.98","Abnormal"]; Sout$p98.sn.wa.h <- cfs.h$cutoffs3["0.98", "Normal"]; 
  Sout$p99.sg.wa.h <- cfs.h$cutoffs3["0.99", "Abnormal"]; Sout$p99.sn.wa.h <- cfs.h$cutoffs3["0.99", "Normal"]; 
}else{
  Sout$p90.sg.w1.h <- Sout$p90.sn.w1.h <- Sout$p95.sg.w1.h <- Sout$p95.sn.w1.h <- Sout$p98.sg.w1.h <- Sout$p98.sn.w1.h <- Sout$p99.sg.w1.h <-  Sout$p99.sn.w1.h <- NA
  Sout$p90.sg.wd.h <- Sout$p90.sn.wd.h <- Sout$p95.sg.wd.h <- Sout$p95.sn.wd.h <- Sout$p98.sg.wd.h <- Sout$p98.sn.wd.h <- Sout$p99.sg.wd.h <-  Sout$p99.sn.wd.h <- NA
  Sout$p90.sg.wa.h <- Sout$p90.sn.wa.h <- Sout$p95.sg.wa.h <- Sout$p95.sn.wa.h <- Sout$p98.sg.wa.h <- Sout$p98.sn.wa.h <- Sout$p99.sg.wa.h <-  Sout$p99.sn.wa.h <- NA
}

write.csv(out, paste0(resdir, Ratio, "_fs_", Score_method, "_", (paste0(ifelse(Lab_curation == "No", "without", "with"), "_LabCureation_")), "E", gsub(", ", "", Exons), "_", params$loessSubset, "_", params$EventCount_Filter, "_MM_", params$MM, "_Outlierremove_",  params$Outlierremove,  "_extrafilter_", extraFilter,".csv"), row.names = FALSE)

write.csv(Sout, paste0(resdir, "Cutoffs_", Ratio, "_fs_", Score_method, "_", (paste0(ifelse(Lab_curation == "No", "without", "with"), "_LabCureation_")), "E", gsub(", ", "", Exons), "_", params$loessSubset, "_", params$EventCount_Filter, "_MM_", params$MM, "_Outlierremove_",  params$Outlierremove,  "_extrafilter_", extraFilter,".csv"), row.names = FALSE)

out2 <- out %>% filter(uPOS %in% D14vsLib_fs.2$uPOS)
write.csv(out2, paste0(resdir, "SNV_", Ratio, "_fs_", Score_method, "_", (paste0(ifelse(Lab_curation == "No", "without", "with"), "_LabCureation_")), "E", gsub(", ", "", Exons), "_", params$loessSubset, "_", params$EventCount_Filter, "_MM_", params$MM, "_Outlierremove_",  params$Outlierremove,  "_extrafilter_", extraFilter,".csv"), row.names = FALSE)

``` 

Output metrics

```{r}
## output the metrics in one file
if(!is.null(params$HDRfile)){
  N_HDR = N_HDR
  AUC_HDR=as.numeric(ROC0.h$auc)
  sens_HDR=ROC0.h$sen
  spec_HDR=ROC0.h$spec
  confusion.matrix_roc_HDR = c(paste(as.vector((c(ROC0.h$tb[1,], ROC0.h$tb[2,]))), collapse = " "))
  
}else{
  N_HDR <- AUC_HDR <- sens_HDR <-  spec_HDR <- NA
  confusion.matrix_roc_HDR = ""
}
                                                   
metrics <- data.frame(Exons = params$Exons,
                      N_SNV = sum(!is.na(out2$functional.score)),
                      EvenFilter = params$EventCount_Filter,
                      vset = params$vset,
                      Method = params$loessSubset,
                      score = Score_method,
                      Ratio = params$Ratio,
                      MM = params$MM,
                      Outlierremove = params$Outlierremove,
                      extraFilter = params$extraFilter,
                      AUC_data=as.numeric(ROC0$auc),
                      sens_data=ROC0$sen,
                      spec_data=ROC0$spec,
                      confusion.matrix_roc_data = c(paste(as.vector((c(ROC0$tb[1,], ROC0$tb[2,]))), collapse = " ")))
if(!is.null(params$clinvarfile)){ 
  metrics <- metrics %>% mutate(
                      AUC_clinvar=as.numeric(ROC0.c$auc),
                      sens_clinvar=ROC0.c$sen,
                      spec_clinvar=ROC0.c$spec,
                      confusion.matrix_roc_clinvar = c(paste(as.vector((c(ROC0.c$tb[1,], ROC0.c$tb[2,]))), collapse = " ")))
}else{
   metrics <- metrics %>% mutate(
                      AUC_clinvar= NA,
                      sens_clinvar=NA,
                      spec_clinvar=NA,
                      confusion.matrix_roc_clinvar = "")
}

metrics <- metrics %>% mutate(
                      N_HDR = N_HDR, 
                      AUC_HDR=AUC_HDR,
                      sens_HDR=sens_HDR,
                      spec_HDR=spec_HDR,
                      confusion.matrix_roc_HDR = confusion.matrix_roc_HDR
                      )


write.csv(metrics, paste0(params$Output_dir, Gene, "_", Ratio, "_", params$vset,"_", params$loessSubset, "_", params$EventCount_Filter, "_", params$Score_method, "_", (paste0(ifelse(Lab_curation == "No", "without", "with"), "_LabCureation_")), "E", gsub(", ", "", Exons), "_MM_", params$MM, "_Outlierremove_",  params$Outlierremove, "_extrafilter_", extraFilter, "_metrics.csv"), row.names = FALSE)

```
